### 2025-04-19 paper/paper.tex
added .tex paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{float} % Add this in your preamble if not included
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Phase 2 - Literature Survey\\}

\author{
\IEEEauthorblockN{Jason Weeks}  
\IEEEauthorblockA{\textit{Department of Computer Science} \\  
\textit{Mississippi State University} \\  
Starkville, Mississippi, USA \\  
jcw1044@msstate.edu}  
\and  
\IEEEauthorblockN{Andrew McBride}  
\IEEEauthorblockA{\textit{Department of Computer Science} \\  
\textit{Mississippi State University} \\  
Starkville, Mississippi, USA \\  
ahm228@msstate.edu}  
\and
\IEEEauthorblockN{Andrei Roskelley Garcia}  
\IEEEauthorblockA{\textit{Department of Computer Science} \\  
\textit{Mississippi State University} \\  
Starkville, Mississippi, USA \\  
ar2888@msstate.edu} 
\and
\IEEEauthorblockN{Jacquies Turner}  
\IEEEauthorblockA{\textit{Department of Computer Science} \\  
\textit{Mississippi State University} \\  
Starkville, Mississippi, USA \\  
jt2485@msstate.edu}  
}

\maketitle

\section{Introduction}
This paper will present the relevant literature in the process of traffic signal control using machine learning techniques, along with the timeline of research developments over time.

\section{Taxonomy}

Research on traffic signal methods falls into three categories. These include single agent learning models, multi agent reinforcement learning models (MARL), and large-scale multi-agent learning models, with the large-scale MARL actively having new deep learning models applied in recent research.

\subsection{Single Agent Learning Models}

Before 2010, traffic prediction models were designed using single agent reinforcement learning models. These models were used to predict the movement of traffic using a single system, meaning a single neural network predicted how traffic lights should function. While innovative, they are generally not scalable for multiple intersections and can more effectively be used to predict how traffic signals function at a single intersection. These models were rarely used in practice and traffic signals used fixed intervals (threshold signaling) \cite{1}. As a result, we will not focus too much on these for our traffic problem.

\subsection{Multi Agent Learning Models}

From 2010-2018, research began to focus on multi-agent learning reinforcement (MARL) models, a more practical approach to traffic signal control. These use several systems, or agents, to predict traffic flow for traffic signal control more accurately, but tend to be limited to knowledge of their own environment and not nearby intersections \cite{2}.

During this research period, the majority of models used a variation of Q-Learning. While not perfect, there were indications that there were improvements in terms of moving traffic over traditional timer methods. \cite{2} Eventually, a distributed MARL signal was designed such that each agent may learn the environment independently and communicate the signals it received to nearby agents to improve performance. This further came with the benefit of allowing one agent to fail without taking other agents down, increasing reliability and scalability \cite{5}. This innovation led to major improvements that have influenced the modern design of reinforcement learning techniques for traffic signal control.

\subsection{Large Scale Multi-Agent Learning Models}

Since 2019, research indicates that MARL agents have been improved for scalability, performance, real-world applications and become open-source. The field has seen enormous growth from the initial small scale solutions, and large scale implementations for real world cities continue to become more practical and powerful.

Early breakthroughs in large-scale MARL focused on a more difficult problem: urban traffic management. This initially came with a new open sourced model, CityFlow, a multi-agent reinforcement model which could handle large road networks with thousands of intersections in a simulated environment \cite{7}. These large-scale models have shown significant improvements in performance while working at a larger scale, making them more adaptable to real situations.

In recent years, strong reinforcement learning models, such as long short-term memory (LSTM) and spatiotemporal graph convolutional neural networks (STGCN) have been applied to research, and external factors like pedestrian movements have begun being addressed. These advancements show more promise in scalability and performance, but are still in progress. There are still opportunities to test different models which may perform better in synchronization with other agents. While accounting for pedestrians marks a major step in optimizing the traffic signal problem, it opens more questions, such as how can we account for emergency vehicles, namely ambulances or police cars, or collisions that disturb the natural flow of traffic. Despite these questions, these models are very powerful in comparison to previous threshold traffic signal management systems.

\subsection{Taxonomy Figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\linewidth]{Research Taxonomy.png}
    \caption{Research Taxonomy for Traffic Signal Control}
\end{figure}

\section{Chronological Overview}
Based on the major research contributions, we have concluded that the papers fall into several categories. These categories tend to reflect the time period of the research contribution, with older papers using less-powerful Q-Learning simulations, and more recent papers using more powerful neural networks with account for more factors. Figure 2 highlights a chronological overview of the categorization of papers over time.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\linewidth]{Diagram Timeline.png}
    \caption{Research Timeline for Traffic Signal Control}
\end{figure}

\section{Research Gaps}
One current research gap is the types of models used. Early MARL models used various Q-learning techniques, with more recent studies using more advanced deep learning techniques, such as LSTM, but transformer-based models and RNN models seem to have been less tested. There are opportunities to test and benchmark different complex deep learning frameworks against each other.

Another research gap is regards real-world application. Several large cities have adapted deep learning techniques for traffic signal processing, though a large majority of cities do not yet have these technologies implemented. Therefore, they could see a change in performance depending on the size of the city and their road systems.

A third research gap appears to be the influence of various real-world distractions that can occur, ranging from road accidents, pedestrians, and emergency vehicles. Although Du et al. have begun to account for pedestrians \cite{13}, there are still several factors that go unaccounted for in the process of traffic signal prediction, which can back up traffic worse than threshold options.

\section*{Work Distribution}
The majority of research findings were made by Jason Weeks and Andrew McBride with help from Andrei by branching out from the original CityFlow paper \cite{7} and finding predecessors and related articles in the field.

Based on the research findings, Jason Weeks, Jacquies Turner, and Andrew McBride contributed to the research taxonomy and timelines. These were made by narrowing down the most important research articles in our original findings.

This paper was primarily written by Jason Weeks, but all the members contributed.

The final distribution of the work is as follows.

\begin{itemize}
    \item 25\% - Jason Weeks
    \item 25\% - Andrew McBride
    \item 25\% - Jacquies Turner
    \item 25\% - Andrei Roskelley Garcia
\end{itemize}

\section{Simulation Environment and Data Preparation}

To effectively evaluate MARL models for traffic signal control, we will employ CityFlow, a multi-agent traffic simulation engine. CityFlow enables efficient traffic modeling by allowing pre-designed scenarios to be simulated instead of relying on real-time data. This approach provides greater flexibility in the definition and testing of traffic environments, making it ideal for research applications. CityFlow is particularly suited for large-scale urban traffic modeling. By simulating traffic flow in a controlled environment, we can test different reinforcement learning models without disrupting real-world traffic. The engine allows for scalable simulations with thousands of intersections, making it a powerful tool for evaluating various traffic control algorithms. For this study, CityFlow will be used to model the traffic network of Starkville, Mississippi, a city that currently employs traditional timer-based traffic light systems. This provides a relevant testing ground for machine learning-based traffic optimization, as improvements in traffic flow could have tangible real-world benefits.

The dataset used for traffic simulation consists of two primary components: roadnet.json and flow.json. The roadnet.json file defines the city's infrastructure, including the road layout, intersections, lane directions, traffic signals, and turning rules. This data is extracted from OpenStreetMaps (OSM) for Starkville and will be converted into a SUMO network before being integrated into CityFlow. Meanwhile, the flow.json file captures the dynamic movement of vehicles within the network, specifying entry and exit points, common routes, and characteristics such as vehicle priority and speed. Since real-world traffic flow data is not readily available, SUMO tools will be used to generate realistic traffic patterns, ensuring an accurate simulation environment to test reinforcement learning-based traffic control strategies.

Starkville presents an ideal case study due to its unique combination of urban and suburban traffic patterns. Unlike many cities that have adopted sensor-based or pressure plate-based signaling systems, Starkville still relies on traditional timer-based traffic signals, making it a prime candidate for testing reinforcement learning-based optimizations.

\section{Implementation}

Our implementation will involve using CityFlow traffic simulation to estimate how traffic flows in Starkville and customize the minutia to our liking much more efficiently than a tool like SUMO. We can collect data from the environment such as congestion levels and states of other environments, along with a set of possible actions to implement a deep Q-learning model, which will reward or punish our model based on actions taken.

Because deep learning is computationally challenging, we will use a joint model. With this model, every action taken will result in a punishment or reward score for the model as a whole. This will allow us to train only one neural network at a time and comes with the added benefit of rewarding actions based on how they affect the state of other intersections. A model like this has historically proven to be more efficient than existing threshold signaling. Our ultimate goal is to prove that this model can provide better traffic signal decisions than existing threshold systems such as those used in Starkville.

\section{Importance and Challenges}

Because traffic light timing plays a crucial role in creating frustrated drivers and accidents, it is critical to find the most effective strategy to create smooth traffic flow through traffic signal control. As a result, our plan for the next two months is to implement a MARL model to predict traffic signals for the city of Starkville Mississippi, a city that uses threshold signaling for various times of the day to move traffic. It could be trained to predict various scenarios, including high-intensity traffic after an athletic event or lighter traffic during school breaks, with the ultimate goal of outperforming existing methodologies for traffic signal processing during all times of the day.

However, there are a variety of challenges with implementation. One challenge is the difficulty in finding optimal Q values as there is no ground truth to compare it to, the agent has to guess and learn on the experience and rewards it receives. This method is also very computationally challenging, as the agent has to constantly update the Q-value, especially with very large and complex models. There is also difficulty in balancing exploration vs exploitation, if the agent explores too much, then the Q-value may not converge to its optimal value, and if it exploits what it knows too much then the value may get stuck at a suboptimal level. Solving this would require understanding and the use of certain exploration strategies. There are also problems that could arise with the environment. The environment can slowly change overtime, so our Q-value may become irrelevant, so the agent would have to relearn its environment. There are also many unexpected situations that may arise that may require the model to quickly learn and adapt in the unknown situation, or try to reinforce the model to account for these situations.

\section{Development plan}

Implementing a traffic simulation for Starkville, MS using CityFlow and Q-Learning over a two-month period, we will begin with the research and environment setup in the first two weeks. This includes installing and configuring CityFlow, generating traffic flows using built-in tools, and setting up a simulation model based on the Starkville road network and intersections. If real-world traffic data is available, we will incorporate it to enhance accuracy; otherwise, default CityFlow scenarios will be used. During weeks three to six, we will develop and train the Q-Learning algorithm by optimizing hyperparameters such as learning rate and exploration rate and running simulations to evaluate its performance. The effectiveness of Q-Learning-controlled signals will be compared against traditional fixed-time signals using metrics like vehicle waiting times and overall traffic flow improvements. In the last two weeks, we will focus on optimizing and validating the model, testing it under different traffic conditions specific to Starkville, Ms. The model will be deployed in a full-scale SUMO simulation, validated against available traffic data, and adjusted accordingly. The project will conclude with documentation of the implementation details and findings, while also identifying opportunities for future improvements, such as deep reinforcement learning for more advanced control. Ultimately, this project aims to develop an intelligent traffic light control system that reduces congestion and improves traffic efficiency in Starkville, MS.

\section{Implementation Details}

\subsection{Simulation}

After beginning implementation we decided that CityFlow was not the appropriate tool for this problem, despite our research. While a powerful tool, it struggled to handle the data format provided from OSM data, resulting in slow performance and broken intersections. We then decided moving to the SUMO simulator was a more worthwhile decision, as its efficency has improved over time and documentation is more appropriate for our specific research area.

\section{ACKNOWLEDGMENT OF LLM USE}

Large language models were not used in the writing or analytical process of writing this paper. However, it was used exclusively to summarize papers during the initial research process, just to ensure that the paper was suitable for our research needs as an aid to abstracts. Not every paper analyzed was used.

\section*{Reference Acknowledgment}
References \cite{1} through \cite{15} highlight the most significant contributions to the study of traffic signal control using machine learning techniques. They highlight how the research has progressed over time and should help guide the development of a model for traffic prediction.

The other sources provide great insight, but might have had a smaller research impact. Despite this, they are still high-quality papers and can be used to help solve our problem. These are still critical to understand the problem at hand, and have a meaningful impact on development of our traffic prediction model.

\begin{thebibliography}{00}
\bibitem{1} Bakker, B., Whiteson, S., Kester, L., \& Groen, F. C. (2010). Traffic light control by multiagent reinforcement learning systems (pp. 475-510). Springer Berlin Heidelberg.
\bibitem{2} Prabuchandran, K. J., AN, H. K., \& Bhatnagar, S. (2014, October). Multi-agent reinforcement learning for traffic signal control. In 17th International IEEE Conference on Intelligent Transportation Systems (ITSC) (pp. 2529-2534). IEEE.
\bibitem{3} Srinivasan, D., Choy, M. C., \& Cheu, R. L. (2006). Neural networks for real-time traffic signal control. IEEE Transactions on intelligent transportation systems, 7(3), 261-272.
\bibitem{4} Xu, M., An, K., Vu, L. H., Ye, Z., Feng, J., \& Chen, E. (2019). Optimizing multi-agent based urban traffic signal control system. Journal of Intelligent Transportation Systems, 23(4), 357-369.
\bibitem{5} Liu, Y., Liu, L., \& Chen, W. P. (2017, October). Intelligent traffic light control using distributed multi-agent Q learning. In 2017 IEEE 20th international conference on intelligent transportation systems (ITSC) (pp. 1-8). IEEE.
\bibitem{6} Cruz-Piris, L., Rivera, D., Fernandez, S., \& Marsa-Maestre, I. (2018). Optimized sensor network and multi-agent decision support for smart traffic light management. Sensors, 18(2), 435.
\bibitem{7} Zhang, H., Feng, S., Liu, C., Ding, Y., Zhu, Y., Zhou, Z., ... \& Li, Z. (2019, May). Cityflow: A multi-agent reinforcement learning environment for large scale city traffic scenario. In The world wide web conference (pp. 3620-3624).
\bibitem{8} Chu, T., Wang, J., Codecà, L., \& Li, Z. (2019). Multi-agent deep reinforcement learning for large-scale traffic signal control. IEEE transactions on intelligent transportation systems, 21(3), 1086-1095.
\bibitem{9} Qadri, S. S. S. M., Gökçe, M. A., \& Öner, E. (2020). State-of-art review of traffic signal control methods: challenges and opportunities. European transport research review, 12, 1-23.
\bibitem{10} Wei, H., Zheng, G., Gayah, V., \& Li, Z. (2021). Recent advances in reinforcement learning for traffic signal control: A survey of models and evaluation. ACM SIGKDD Explorations Newsletter, 22(2), 12-18.
\bibitem{11} Abdoos, M., \& Bazzan, A. L. (2021). Hierarchical traffic signal optimization using reinforcement learning and traffic prediction with long-short term memory. Expert systems with applications, 171, 114580.
\bibitem{12} Kolat, M., Kővári, B., Bécsi, T., \& Aradi, S. (2023). Multi-agent reinforcement learning for traffic signal control: A cooperative approach. Sustainability, 15(4), 3479.
\bibitem{13} Du, W., Ye, J., Gu, J., Li, J., Wei, H., \& Wang, G. (2023, June). Safelight: A reinforcement learning method toward collision-free traffic signal control. In Proceedings of the AAAI conference on artificial intelligence (Vol. 37, No. 12, pp. 14801-14810).
\bibitem{14} Yazdani, M., Sarvi, M., Bagloee, S. A., Nassir, N., Price, J., \& Parineh, H. (2023). Intelligent vehicle pedestrian light (IVPL): A deep reinforcement learning approach for traffic signal control. Transportation research part C: emerging technologies, 149, 103991.
\bibitem{15} Lin, T., \& Lin, R. (2024). Smart City Traffic Flow and Signal Optimization Using STGCN-LSTM and PPO Algorithms. IEEE Access.
\bibitem{16} Haydari, A., \& Yılmaz, Y. (2020). Deep reinforcement learning for intelligent transportation systems: A survey. IEEE Transactions on Intelligent Transportation Systems, 23(1), 11-32.
\bibitem{17} Wei, H., Xu, N., Zhang, H., Zheng, G., Zang, X., Chen, C., ... \& Li, Z. (2019, November). Colight: Learning network-level cooperation for traffic signal control. In Proceedings of the 28th ACM international conference on information and knowledge management (pp. 1913-1922).
\bibitem{18} Chen, C., Wei, H., Xu, N., Zheng, G., Yang, M., Xiong, Y., ... \& Li, Z. (2020, April). Toward a thousand lights: Decentralized deep reinforcement learning for large-scale traffic signal control. In Proceedings of the AAAI conference on artificial intelligence (Vol. 34, No. 04, pp. 3414-3421).
\bibitem{19} Zheng, G., Xiong, Y., Zang, X., Feng, J., Wei, H., Zhang, H., ... \& Li, Z. (2019, November). Learning phase competition for traffic signal control. In Proceedings of the 28th ACM international conference on information and knowledge management (pp. 1963-1972).
\bibitem{20} Wei, H., Zheng, G., Gayah, V., \& Li, Z. (2019). A survey on traffic signal control methods. arXiv preprint arXiv:1904.08117.
\bibitem{21} Wei, H., Zheng, G., Yao, H., \& Li, Z. (2018, July). Intellilight: A reinforcement learning approach for intelligent traffic light control. In Proceedings of the 24th ACM SIGKDD international conference on knowledge discovery \& data mining (pp. 2496-2505).
\bibitem{22} Yang, L., Luo, P., Change Loy, C., \& Tang, X. (2015). A large-scale car dataset for fine-grained categorization and verification. In Proceedings of the IEEE conference on computer vision and pattern recognition (pp. 3973-3981).
\bibitem{23} Milan, A. (2016). MOT16: A benchmark for multi-object tracking. arXiv preprint arXiv:1603.00831.
\bibitem{24} Mao, F., Li, Z., \& Li, L. (2022). A comparison of deep reinforcement learning models for isolated traffic signal control. IEEE Intelligent Transportation Systems Magazine, 15(1), 160-180.
\bibitem{25} Noaeen, M., Naik, A., Goodman, L., Crebo, J., Abrar, T., Abad, Z. S. H., ... \& Far, B. (2022). Reinforcement learning in urban network traffic signal control: A systematic literature review. Expert Systems with Applications, 199, 116830.
\bibitem{26} Claus, C., \& Boutilier, C. (1998). The dynamics of reinforcement learning in cooperative multiagent systems. AAAI/IAAI, 1998(746-752), 2.
\bibitem{27} Arel, I., Liu, C., Urbanik, T., \& Kohls, A. G. (2010). Reinforcement learning-based multi-agent system for network traffic signal control. IET Intelligent Transport Systems, 4(2), 128-135.
\bibitem{28} Jácome, L., Benavides, L., Jara, D., Riofrio, G., Alvarado, F., \& Pesantez, M. (2018, October). A survey on intelligent traffic lights. In 2018 IEEE International Conference on Automation/XXIII Congress of the Chilean Association of Automatic Control (ICA-ACCA) (pp. 1-6). IEEE.
\bibitem{29} Mei, H., Lei, X., Da, L., Shi, B., \& Wei, H. (2024). Libsignal: an open library for traffic signal control. Machine Learning, 113(8), 5235-5271.
\bibitem{30} Eom, M., \& Kim, B. I. (2020). The traffic signal control problem for intersections: a review. European transport research review, 12, 1-20.
\bibitem{31} El-Tantawy, S., \& Abdulhai, B. (2010, September). An agent-based learning towards decentralized and coordinated traffic signal control. In 13th International IEEE conference on intelligent transportation systems (pp. 665-670). IEEE.
\bibitem{32} Tang, Z., Naphade, M., Liu, M. Y., Yang, X., Birchfield, S., Wang, S., ... \& Hwang, J. N. (2019). Cityflow: A city-scale benchmark for multi-target multi-camera vehicle tracking and re-identification. In Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition (pp. 8797-8806).
\bibitem{33} Cools, S. B., Gershenson, C., \& D’Hooghe, B. (2013). Self-organizing traffic lights: A realistic simulation. Advances in applied self-organizing systems, 45-55.
\bibitem{34} Bouktif, S., Cheniki, A., Ouni, A., \& El-Sayed, H. (2023). Deep reinforcement learning for traffic signal control with consistent state and reward design approach. Knowledge-Based Systems, 267, 110440.
\bibitem{35} Ducrocq, R., \& Farhi, N. (2023). Deep reinforcement Q-learning for intelligent traffic signal control with partial detection. International journal of intelligent transportation systems research, 21(1), 192-206.
\bibitem{36} Wang, X., Ke, L., Qiao, Z., \& Chai, X. (2020). Large-scale traffic signal control using a novel multiagent reinforcement learning. IEEE transactions on cybernetics, 51(1), 174-187.
\bibitem{37} Wei, H., Chen, C., Zheng, G., Wu, K., Gayah, V., Xu, K., \& Li, Z. (2019, July). Presslight: Learning max pressure control to coordinate traffic signals in arterial network. In Proceedings of the 25th ACM SIGKDD international conference on knowledge discovery \& data mining (pp. 1290-1298).
\bibitem{38} Zang, X., Yao, H., Zheng, G., Xu, N., Xu, K., \& Li, Z. (2020, April). Metalight: Value-based meta-reinforcement learning for traffic signal control. In Proceedings of the AAAI conference on artificial intelligence (Vol. 34, No. 01, pp. 1153-1160).
\bibitem{39} Zhang, H., Liu, C., Zhang, W., Zheng, G., \& Yu, Y. (2020, October). Generalight: Improving environment generalization of traffic signal control via meta reinforcement learning. In Proceedings of the 29th ACM international conference on information \& knowledge management (pp. 1783-1792).
\bibitem{40} Navarro-Espinoza, A., López-Bonilla, O. R., García-Guerrero, E. E., Tlelo-Cuautle, E., López-Mancilla, D., Hernández-Mejía, C., \& Inzunza-González, E. (2022). Traffic flow prediction for smart traffic lights using machine learning algorithms. Technologies, 10(1), 5.
\bibitem{41} Natafgi, M. B., Osman, M., Haidar, A. S., \& Hamandi, L. (2018, November). Smart traffic light system using machine learning. In 2018 IEEE International Multidisciplinary Conference on Engineering Technology (IMCET) (pp. 1-6). IEEE.
\bibitem{42} Behrendt, K., Novak, L., \& Botros, R. (2017, May). A deep learning approach to traffic lights: Detection, tracking, and classification. In 2017 IEEE International Conference on Robotics and Automation (ICRA) (pp. 1370-1377). IEEE.
\bibitem{43} Ottom, M. A., \& Al-Omari, A. (2023). An adaptive traffic lights system using machine learning. Int. Arab J. Inf. Technol., 20(3), 407-418.
\bibitem{44} Kumar, N., Mittal, S., Garg, V., \& Kumar, N. (2021). Deep reinforcement learning-based traffic light scheduling framework for SDN-enabled smart transportation system. IEEE Transactions on Intelligent Transportation Systems, 23(3), 2411-2421.
\bibitem{45} Sagiroglu, S., Yavanoglu, U., \& Guven, E. N. (2007, December). Web based machine learning for language identification and translation. In Sixth International Conference on Machine Learning and Applications (ICMLA 2007) (pp. 280-285). IEEE.
\bibitem{46} Li, Y., Zhang, Y., Li, X., \& Sun, C. (2024). Regional multi-agent cooperative reinforcement learning for city-level traffic grid signal control. IEEE/CAA Journal of Automatica Sinica, 11(9), 1987-1998.
\bibitem{47} Peng, X., Gao, H., Han, G., Wang, H., \& Zhang, M. (2023). Joint Optimization of Traffic Signal Control and Vehicle Routing in Signalized Road Networks using Multi-Agent Deep Reinforcement Learning. arXiv preprint arXiv:2310.10856.
\bibitem{48} Zhang, Y., Yu, Z., Zhang, J., Wang, L., Luan, T. H., Guo, B., \& Yuen, C. (2023). Learning Decentralized Traffic Signal Controllers with Multi-Agent Graph Reinforcement Learning. IEEE Transactions on Mobile Computing.
\bibitem{49} Wang, K., Shen, Z., Lei, Z., \& Zhang, T. (2024). Towards Multi-agent Reinforcement Learning based Traffic Signal Control through Spatio-temporal Hypergraphs. arXiv preprint arXiv:2404.11014.
\bibitem{50} Chen, F., Chen, Z., Biswas, S., Lei, S., Ramakrishnan, N., \& Lu, C. T. (2020, November). Graph convolutional networks with kalman filtering for traffic prediction. In Proceedings of the 28th international conference on advances in geographic information systems (pp. 135-138).
\bibitem{51} Li, Z., Xu, C., \& Zhang, G. (2021). A deep reinforcement learning approach for traffic signal control optimization. arXiv preprint arXiv:2107.06115.
\bibitem{52} Zhang, Y., Zheng, G., Liu, Z., Li, Q., \& Zeng, H. (2024). MARLens: understanding multi-agent reinforcement learning for traffic signal control via visual analytics. IEEE transactions on visualization and computer graphics.
\end{thebibliography}

\end{document}

### 2025-04-19 config.py
Update config.py
Justifications:
NUM_INTERSECTIONS = 62, number of instances of 'traffic_light' in map.xml
STATE_DIM = 312, 62 * 5. 5 features per lighted intersection. queue length, max wait time in queue, avg wait time in queue, etc.
ACTIONS_PER_INT = 3, red, yellow, green. Turn signals handled seperately.
GAMMA γ=0.01^(1/300).  1s simulation step, 300 second simulation time. Adjust as needed.
BATCH_SIZE may need increasing to 256 or 512
LEARNING_RATE should be good as is.
TARGET_UPDATE_FREQUENCY should be good as is
REPLAY_BUFFER_CAPACITY increased to properly handle 15 episodes at 1s per s

@@ -25,14 +25,15 @@ CONTROL_INTERVAL = 5       # use the same action for 5 simulation steps to avoid
 
 #TODO: Fine tune these @Andrei
 # DQN hyperparameters 
-STATE_DIM = 1000           # total dimension of state vector (for example, 100 intersections * 10 features each)
-ACTIONS_PER_INT = 4        # number of phases per intersection (assuming all intersections have the same count)
-NUM_INTERSECTIONS = 100    # total intersections controlled
+NUM_INTERSECTIONS = 62    # total intersections controlled
+FEATURES_PER_INT = 5
+STATE_DIM = NUM_INTERSECTIONS * FEATURES_PER_INT          # total dimension of state vector (for example, 100 intersections * 10 features each)
+ACTIONS_PER_INT = 3        # number of phases per intersection (assuming all intersections have the same count)
 BATCH_SIZE = 64
-GAMMA = 0.99
+GAMMA = 0.01**(1/300)       # γ^H=δ ⟹ γ=δ^(1/H)
 LEARNING_RATE = 1e-4
 TARGET_UPDATE_FREQ = 1000  # steps
-REPLAY_BUFFER_CAPACITY = 100000
+REPLAY_BUFFER_CAPACITY = 100800
 
 # Epsilon parameters for epsilon-greedy exploration
 EPS_START = 1.0

### 2025-04-19 config.py
Update config.py

@@ -29,7 +29,7 @@ NUM_INTERSECTIONS = 62    # total intersections controlled
 FEATURES_PER_INT = 5
 STATE_DIM = NUM_INTERSECTIONS * FEATURES_PER_INT          # total dimension of state vector (for example, 100 intersections * 10 features each)
 ACTIONS_PER_INT = 3        # number of phases per intersection (assuming all intersections have the same count)
-BATCH_SIZE = 64
+BATCH_SIZE = 64     # increase if updates are too noisy
 GAMMA = 0.01**(1/300)       # γ^H=δ ⟹ γ=δ^(1/H)
 LEARNING_RATE = 1e-4
 TARGET_UPDATE_FREQ = 1000  # steps
@@ -38,4 +38,4 @@ REPLAY_BUFFER_CAPACITY = 100800
 # Epsilon parameters for epsilon-greedy exploration
 EPS_START = 1.0
 EPS_END = 0.05
-EPS_DECAY = 5000  # decay steps
+EPS_DECAY = 5000  # decay steps  #NOTE: slow it down if the agent “goes greedy” before it’s learned anything

### 2025-04-19 env.py
get_state
Define and extract features

@@ -3,7 +3,7 @@
 import traci
 import random
 import numpy as np
-from config import SUMO_START_CMD, SUMO_LOAD_CMD, CONTROL_INTERVAL, MAX_STEPS
+from config import SUMO_START_CMD, SUMO_LOAD_CMD, CONTROL_INTERVAL, MAX_STEPS, FEATURES_PER_INT
 
 class SUMOEnv:
     def __init__(self):
@@ -61,19 +61,40 @@ class SUMOEnv:
         return next_state, cumulative_reward, done
     
     def get_state(self):
-        """
-        Gather the current state from SUMO.
-        For example, extract the current phase of each traffic light and dummy features.
-        Replace the dummy features with your actual sensor data.
-        """
         state = []
         for tls in self.tls_ids:
+            # 1) Current phase index
             phase = traci.trafficlight.getPhase(tls)
-            # Dummy features: normalize the phase and add random numbers for demonstration.
-            features = [phase / 10.0]
-            # Example: extend with other 9 dummy features.
-            features.extend([random.random() for _ in range(9)])
+
+            # 2) Queue length on inbound edge
+            inflows = traci.trafficlight.getControlledLinks(tls)
+            # flatten link list to edge IDs
+            edges = [link[0] for group in inflows for link in group]
+            queue_lens = [traci.edge.getLastStepHaltingNumber(e) for e in edges]
+
+            # 3) Total waiting time on those edges
+            wait_times = []
+            for e in edges:
+                for vid in traci.edge.getLastStepVehicleIDs(e):
+                    wait_times.append(traci.vehicle.getWaitingTime(vid))
+
+            # 4) Instantaneous delay estimate (timeLoss is cumulative, so use speed)
+            speeds = [traci.edge.getLastStepMeanSpeed(e) for e in edges]
+            freespeed = [traci.edge.getMaxSpeed(e) for e in edges]
+            # normalized delay per edge 1 - (speed/maxSpeed)
+            delays = [1 - s/ms for s,ms in zip(speeds, freespeed)]
+
+            features = [
+                phase / len(self.tls_ids),  # normalized phase
+                sum(queue_lens),            # total queue
+                sum(wait_times),            # total wait
+                np.mean(delays),            # average delay
+                len(edges)                  # number of inbound links
+            ]
+            assert len(features) == FEATURES_PER_INT
+
             state.extend(features)
+            
         return np.array(state, dtype=np.float32)
     
     def compute_reward(self):

### 2025-04-19 env.py
Filled out compute_reward

@@ -1,7 +1,6 @@
 # env.py
 
 import traci
-import random
 import numpy as np
 from config import SUMO_START_CMD, SUMO_LOAD_CMD, CONTROL_INTERVAL, MAX_STEPS, FEATURES_PER_INT
 
@@ -94,25 +93,38 @@ class SUMOEnv:
             assert len(features) == FEATURES_PER_INT
 
             state.extend(features)
-            
+
         return np.array(state, dtype=np.float32)
     
     def compute_reward(self):
-        """
-        Compute reward based on your simulation metrics.
-        Here we use a dummy reward, but you can replace this with your own logic.
-        """
+
+        total_queue = 0.0
         total_wait = 0.0
-        max_q = 0
+        total_delay = 0.0
         
-        # If you have defined monitored_edges, compute waiting times and queue lengths.
-        for edge in self.monitored_edges:
-            q = traci.edge.getLastStepHaltingNumber(edge)
-            max_q = max(max_q, q)
-            veh_ids = traci.edge.getLastStepVehicleIDs(edge)
-            for veh in veh_ids:
-                total_wait += traci.vehicle.getWaitingTime(veh)
-        reward = - (total_wait + 5 * max_q)
+        # Loop over each traffic light exactly as in get_state()
+        for tls in self.tls_ids:
+            # 1) gather the controlled edges
+            inflows = traci.trafficlight.getControlledLinks(tls)
+            edges = [link[0] for group in inflows for link in group]
+            
+            # 2) queue lengths
+            qlens = [traci.edge.getLastStepHaltingNumber(e) for e in edges]
+            total_queue += sum(qlens)
+            
+            # 3) waiting times
+            for e in edges:
+                for vid in traci.edge.getLastStepVehicleIDs(e):
+                    total_wait += traci.vehicle.getWaitingTime(vid)
+            
+            # 4) instantaneous delay estimate
+            speeds    = [traci.edge.getLastStepMeanSpeed(e) for e in edges]
+            freespeed = [traci.edge.getMaxSpeed(e)         for e in edges]
+            # sum of (1 - speed/maxSpeed) over all edges
+            total_delay += sum((1 - s/ms) for s,ms in zip(speeds, freespeed) if ms>0)
+        
+        reward = - ( total_queue + total_wait + 0.5 * total_delay )
+        
         return reward
     
     def close(self):

### 2025-04-20 paper/paper.tex
Init
initalize new files

@@
 The project will conclude with documentation of the implementation details and findings, while also identifying opportunities for future improvements, such as deep reinforcement learning for more advanced control.
-Ultimately, this project aims to develop an intelligent traffic light control system that reduces congestion and improves traffic efficiency in Starkville, MS.
+Ultimately, this project aims to develop an intelligent traffic light control system that reduces congestion and improves traffic efficiency in Starkville, Mississippi.
 
-\section{Implementation Details}
+\section{Implemented Application}
 
-\subsection{Simulation}
+\subsection{Simulation Setup}
 
@@
 We then decided moving to the SUMO simulator was a more worthwhile decision, as its efficency has improved over time and documentation is more appropriate for our specific research area.
+
+Once SUMO was installed, we downloaded public data for the city of Starkville Mississippi's road structure using the tool open street map, and used SUMO's netedit to process the data into a useable format.
+Some changes were made, such as fixing traffic signal placement using junctions, and fixing non-signal based intersection interactions.
+
+Our ultimate goal with this project is to test how neural network influenced traffic signals direct the flow of traffic, so we needed to setup various simulation environments to simulate realistic traffic scenarios within Starkville.
+To do this, we used SUMO's built in flow generator, with various parameters to result in light, medium, and intense traffic scenarios.
+This would allow us to test how each method of traffic signal control reacts to various events such as morning and afternoon rush hours, post game traffic, or light summer traffic.
+
+Using the setup configuration file, we can use SUMO GUI, which comes bundled with the installation package of SUMO.
+This tools allows us to run a visualization of the simulation with regards to the currently implemented method of traffic signal control, and make observations from the point of view of drivers without quantitative analysis, making observations about our model to make improvements.
+
+**INSERT IMAGE OF SUMO RUNNING**
+
+\subsection{Benchmarking}
+
+In order to properly evaluate the performance of threshold and neural network models, a method of evaluation was needed beyond qualitative methods, such as observation or reivew.
+As a result we implemented various benchmarks to quantitatively decide the quality of a methods approach.
+
+Our first method of benchmark involves calculating the average waiting time of a vehicle. This is the most likely representation of how frustrated one might feel.
+
+The next is the average trip time
+
+Finally we calculated information about queue length.
+
+\subsection{Neural Network Implementation}
+
+\section{Results over baseline}
 

### 2025-04-20 agent.py

@@ -1,95 +1,0 @@
-# agent.py
-
-import torch
-import torch.optim as optim
-import torch.nn.functional as F
-import numpy as np
-from model import CityDQN
-from replay_buffer import ReplayBuffer
-from config import (STATE_DIM, ACTIONS_PER_INT, NUM_INTERSECTIONS, 
-                    GAMMA, LEARNING_RATE, TARGET_UPDATE_FREQ, 
-                    REPLAY_BUFFER_CAPACITY, BATCH_SIZE, EPS_START, EPS_END, EPS_DECAY)
-
-class DQNAgent:
-    def __init__(self, device):
-        self.device = device
-        self.state_dim = STATE_DIM
-        self.actions_per_int = ACTIONS_PER_INT
-        self.num_intersections = NUM_INTERSECTIONS
-
-        self.policy_net = CityDQN(self.state_dim, self.actions_per_int, self.num_intersections).to(self.device)
-        self.target_net = CityDQN(self.state_dim, self.actions_per_int, self.num_intersections).to(self.device)
-        self.target_net.load_state_dict(self.policy_net.state_dict())
-        self.target_net.eval()
-
-        self.optimizer = optim.Adam(self.policy_net.parameters(), lr=LEARNING_RATE)
-        self.replay_buffer = ReplayBuffer(REPLAY_BUFFER_CAPACITY)
-        self.steps_done = 0
-
-    def select_action(self, state):
-        """
-        Select a joint action (a list with one action per intersection) using epsilon-greedy strategy.
-        """
-        sample = np.random.rand()
-        eps_threshold = EPS_END + (EPS_START - EPS_END) * np.exp(-1. * self.steps_done / EPS_DECAY)
-        self.steps_done += 1
-        if sample < eps_threshold:
-            # Random action for each intersection
-            return [np.random.randint(self.actions_per_int) for _ in range(self.num_intersections)]
-        else:
-            # Use policy_net
-            with torch.no_grad():
-                state_tensor = torch.tensor(state, dtype=torch.float32).unsqueeze(0).to(self.device)
-                q_values = self.policy_net(state_tensor)
-                # Reshape to (num_intersections, actions_per_int)
-                q_values = q_values.view(self.num_intersections, self.actions_per_int)
-                # Choose the best action per intersection
-                actions = q_values.argmax(dim=1)
-                return actions.cpu().numpy().tolist()
-
-    def train_step(self):
-        if len(self.replay_buffer) < BATCH_SIZE:
[…]

### 2025-04-20 env.py

@@ -1,98 +1,0 @@
-# env.py
-
-import traci
-import random
-import numpy as np
-from config import SUMO_START_CMD, SUMO_LOAD_CMD, CONTROL_INTERVAL, MAX_STEPS
-
-class SUMOEnv:
-    def __init__(self):
-        # Use the start command for starting SUMO (which includes the full executable path)
-        # and the load command for reloading.
-        self.start_cmd = SUMO_START_CMD
-        self.load_cmd = SUMO_LOAD_CMD
-        self.sim_step = 0
-        self.tls_ids = None  # Will be populated when starting SUMO.
-        self.monitored_edges = []  # Populate with your edge IDs or leave empty for now.
-    
-    def start(self):
-        # Start the SUMO simulation.
-        traci.start(self.start_cmd)
-        self.tls_ids = traci.trafficlight.getIDList()
-    
-    def reset(self):
-        # Reload the simulation using parameters allowed in traci.load()
-        # Option 1: Use traci.load() if you want a "soft reset":
-        traci.load(self.load_cmd)
-        self.sim_step = 0
-        return self.get_state()
-        
-        # Option 2 (Alternate): Restart the simulation completely:
-        # traci.close()
-        # traci.start(self.start_cmd)
-        # self.sim_step = 0
-        # return self.get_state()
-    
-    def step(self, joint_action):
-        """
-        Apply a joint action, which is a list of actions for each traffic light.
-        The method simulates CONTROL_INTERVAL steps with that action.
-        Returns: next_state, cumulative_reward, done flag.
-        """
-        cumulative_reward = 0.0
-        
-        # Execute the same action for CONTROL_INTERVAL simulation steps.
-        for _ in range(CONTROL_INTERVAL):
-            for idx, tls in enumerate(self.tls_ids):
-                # Apply the action for each traffic light using the given phase index.
-                # Adjust this call as needed (e.g., if you use setRedYellowGreenState instead)
-                traci.trafficlight.setPhase(tls, int(joint_action[idx]))
-            traci.simulationStep()
-            self.sim_step += 1
-            cumulative_reward += self.compute_reward()
-            
-            # Check termination condition (e.g., max simulation steps reached)
-            if self.sim_step >= MAX_STEPS:
-                next_state = self.get_state()
-                return next_state, cumulative_reward, True
-                
-        next_state = self.get_state()
-        done = False
-        return next_state, cumulative_reward, done
-    
-    def get_state(self):
-        """
-        Gather the current state from SUMO.
[…]

### 2025-04-20 main.py

@@ -1,52 +1,0 @@
-# main.py
-
-import time
-import torch
-import numpy as np
-from env import SUMOEnv
-from agent import DQNAgent
-from config import NUM_EPISODES, TARGET_UPDATE_FREQ
-
-def main():
-    # Use GPU if available; otherwise CPU.
-    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
-    print("Using device:", device)
-
-    # Initialize environment and agent.
-    env = SUMOEnv()
-    env.start()
-    agent = DQNAgent(device)
-
-    global_step = 0
-
-    for episode in range(NUM_EPISODES):
-        state = env.reset()
-        episode_reward = 0
-        done = False
-
-        while not done:
-            # Select joint action using epsilon-greedy.
-            action = agent.select_action(state)
-            # Step the simulation for CONTROL_INTERVAL steps and accumulate reward.
-            next_state, reward, done = env.step(action)
-            episode_reward += reward
-
-            # Add transition to replay buffer.
-            agent.replay_buffer.add(state, action, reward, next_state, done)
-            state = next_state
-
-            # Train the agent at every step.
-            loss = agent.train_step()
-            global_step += 1
-
-            # Update target network periodically.
-            if global_step % TARGET_UPDATE_FREQ == 0:
-                agent.update_target()
-                print(f"Updated target network at step {global_step}")
-
-        print(f"Episode {episode+1}: Total Reward = {episode_reward}")
-    
-    env.close()
-
-if __name__ == '__main__':
-    main()

### 2025-04-20 model.py

@@ -1,23 +1,0 @@
-# model.py
-
-import torch
-import torch.nn as nn
-
-class CityDQN(nn.Module):
-    def __init__(self, state_dim, actions_per_int, num_intersections):
-        super(CityDQN, self).__init__()
-        self.state_dim = state_dim
-        self.actions_per_int = actions_per_int
-        self.num_intersections = num_intersections
-        # The total output is factorized: each intersection gets its own set of Q-values.
-        self.action_dim = actions_per_int * num_intersections
-
-        # A simple network structure – adjust number of neurons and layers as needed.
-        self.fc1 = nn.Linear(state_dim, 256)
-        self.fc2 = nn.Linear(256, 256)
-        self.out = nn.Linear(256, self.action_dim)
-
-    def forward(self, x):
-        x = torch.relu(self.fc1(x))
-        x = torch.relu(self.fc2(x))
-        return self.out(x)  # flat vector; later reshaped to (num_intersections, actions_per_int)

### 2025-04-20 traffic_signal_nn/agents/models.py

@@ -1,0 +1,16 @@
+# traffic_signal_nn/agents/models.py
+
+import torch.nn as nn
+
+def build_mlp(input_dim, output_dim, hidden_units):
+    """
+    Construct a plain MLP: FC → ReLU → … → FC.
+    """
+    layers = []
+    prev = input_dim
+    for h in hidden_units:
+        layers.append(nn.Linear(prev, h))
+        layers.append(nn.ReLU())
+        prev = h
+    layers.append(nn.Linear(prev, output_dim))
+    return nn.Sequential(*layers)

### 2025-04-20 traffic_signal_nn/agents/policies.py

@@ -1,0 +1,85 @@
+# traffic_signal_nn/agents/policies.py
+
+import random
+import torch
+import torch.optim as optim
+import torch.nn.functional as F
+from .models import build_mlp
+from .replay_buffer import ReplayBuffer
+
+def get_agent_class(name):
+    name = name.lower()
+    if name == "dqn":
+        return DQNAgent
+    raise ValueError(f"Unknown agent '{name}'")
+
+class DQNAgent:
+    """
+    Standard (single‑agent) DQN.
+    """
+    def __init__(self, state_dim, action_dim, cfg):
+        self.state_dim  = state_dim
+        self.action_dim = action_dim
+        self.lr         = cfg["LR"]
+        self.gamma      = cfg["GAMMA"]
+        self.epsilon    = cfg["EPS_START"]
+        self.eps_end    = cfg["EPS_END"]
+        self.eps_decay  = cfg["EPS_DECAY"]
+
+        hidden = list(map(int, cfg["HIDDEN"].split(",")))
+        self.policy_net = build_mlp(state_dim, action_dim, hidden)
+        self.target_net = build_mlp(state_dim, action_dim, hidden)
+        self.target_net.load_state_dict(self.policy_net.state_dict())
+
+        self.optimizer   = optim.Adam(self.policy_net.parameters(), lr=self.lr)
+        self.replay_buf  = ReplayBuffer(int(cfg["BUFFER_SIZE"]))
+        self.batch_size  = cfg["BATCH_SIZE"]
+        self.update_freq = cfg.get("TARGET_UPDATE_FREQ", 10)
+        self.step_count  = 0
+        self.loss        = 0.0
+
+    def select_action(self, state):
+        if random.random() < self.epsilon:
+            return random.randrange(self.action_dim)
+        with torch.no_grad():
+            qvals = self.policy_net(torch.tensor(state, dtype=torch.float32))
+            return int(qvals.argmax().item())
+
+    def remember(self, s, a, r, ns, done):
+        self.replay_buf.add(s, a, r, ns, done)
+
+    def learn(self):
+        if len(self.replay_buf) < self.batch_size:
+            return
+
+        states, actions, rewards, next_states, dones = self.replay_buf.sample(self.batch_size)
+        s  = torch.tensor(states, dtype=torch.float32)
+        a  = torch.tensor(actions, dtype=torch.long).unsqueeze(1)
+        r  = torch.tensor(rewards, dtype=torch.float32).unsqueeze(1)
[…]

### 2025-04-20 traffic_signal_nn/agents/replay_buffer.py

@@ -1,11 +1,13 @@
-# replay_buffer.py
+# traffic_signal_nn/agents/replay_buffer.py
 
 import random
 from collections import deque
 
 class ReplayBuffer:
+    """
+    Fixed‑size FIFO buffer for experience replay.
+    """
     def __init__(self, capacity):
-        self.capacity = capacity
         self.buffer = deque(maxlen=capacity)
 
     def add(self, state, action, reward, next_state, done):
@@ -13,7 +15,8 @@ class ReplayBuffer:
 
     def sample(self, batch_size):
         batch = random.sample(self.buffer, batch_size)
-        return batch
+        states, actions, rewards, next_states, dones = zip(*batch)
+        return list(states), list(actions), list(rewards), list(next_states), list(dones)
 
     def __len__(self):
         return len(self.buffer)

### 2025-04-20 traffic_signal_nn/env/base_env.py

@@ -1,0 +1,55 @@
+# traffic_signal_nn/env/base_env.py
+
+import traci
+
+class BaseEnv:
+    """
+    Wrap a single‑junction SUMO sim as an OpenAI‑style env.
+    """
+    def __init__(self, cfg):
+        # cfg should contain SUMO_BINARY, SUMO_CFG, STEP_LEN
+        self.sumo_binary = cfg.get("SUMO_BINARY", "sumo")
+        self.sumocfg     = cfg["SUMO_CFG"]
+        self.step_len    = cfg.get("STEP_LEN", 1)
+        self.sumo_cmd    = [self.sumo_binary, "-c", self.sumocfg]
+        self._start_sumo()
+        self.lanes       = traci.lane.getIDList()
+
+    def _start_sumo(self):
+        traci.start(self.sumo_cmd)
+
+    def reset(self):
+        traci.load(self.sumo_cmd)
+        self._start_sumo()
+        return self.get_state()
+
+    def step(self, action):
+        self._apply_action(action)
+        traci.simulationStep()
+        next_state = self.get_state()
+        reward     = self._compute_reward()
+        done       = traci.simulation.getMinExpectedNumber() == 0
+        return next_state, reward, done, {}
+
+    def _apply_action(self, action):
+        """
+        Must be overridden: map `action` to traffic‑light phase changes.
+        """
+        raise NotImplementedError
+
+    def _compute_reward(self):
+        """
+        Default reward = –total queue.
+        """
+        return -sum(traci.lane.getLastStepHaltingNumber(l)
+                    for l in self.lanes)
+
+    def get_state(self):
+        """
+        Default state = vector of per‑lane queue lengths.
+        """
+        return [traci.lane.getLastStepHaltingNumber(lane)
+                for lane in self.lanes]
+
+    def close(self):
+        traci.close()

### 2025-04-20 traffic_signal_nn/env/city_env.py

@@ -1,0 +1,32 @@
+# traffic_signal_nn/env/city_env.py
+
+from .base_env import BaseEnv
+import traci
+
+class CityEnv(BaseEnv):
+    """
+    A single‑junction env for your target city.
+    """
+    def __init__(self, cfg):
+        """
+        cfg must include:
+          - SUMO_CFG: path to simulation.sumocfg
+          - TLS_ID:  traffic‑light system ID in the network
+          - N_PHASES: number of signal phases
+        """
+        super().__init__(cfg)
+        self.tls_id   = cfg["TLS_ID"]
+        self.n_phases = cfg["N_PHASES"]
+
+    def _apply_action(self, action):
+        # assume `action` is int phase index
+        traci.trafficlight.setPhase(self.tls_id, int(action))
+
+    @property
+    def observation_space(self):
+        # simple vector length = #lanes
+        return len(self.lanes)
+
+    @property
+    def action_space(self):
+        return self.n_phases

### 2025-04-20 traffic_signal_nn/main.py

@@ -1,0 +1,67 @@
+import argparse
+import os
+from utils.config_parser import load_config
+from utils.logger import Logger
+from env.city_env import CityEnv
+from agents.policies import get_agent_class
+
+def main():
+    parser = argparse.ArgumentParser(description="Multi‑RL Traffic Signal Control")
+    parser.add_argument("--config",     required=True, help="Path to .ini config")
+    parser.add_argument("mode",         choices=["train","eval","demo"],
+                        help="train: learn; eval: run saved model; demo: render only")
+    args = parser.parse_args()
+
+    # load all sections into nested dicts
+    cfg = load_config(args.config)
+
+    # make env & agent
+    env_cls = CityEnv
+    env     = env_cls(cfg["ENV"])
+    Agent   = get_agent_class(cfg["AGENT"]["NAME"])
+    agent   = Agent(env.observation_space,
+                    env.action_space,
+                    cfg["AGENT"])
+
+    # make logger
+    logger = Logger(cfg["LOG"]["LOG_DIR"])
+
+    if args.mode == "train":
+        for episode in range(cfg["TRAIN"]["EPISODES"]):
+            state, done, ep_reward = env.reset(), False, 0.0
+            while not done:
+                action = agent.select_action(state)
+                next_state, reward, done, _ = env.step(action)
+                agent.remember(state, action, reward, next_state, done)
+                agent.learn()
+                state, ep_reward = next_state, ep_reward + reward
+
+            agent.update_target_network()
+            logger.log_episode(episode, ep_reward, agent)
+
+        os.makedirs(logger.log_dir, exist_ok=True)
+        agent.save(os.path.join(logger.log_dir, "agent.pth"))
+
+    elif args.mode == "eval":
+        agent.load(os.path.join(cfg["LOG"]["LOAD_DIR"], "agent.pth"))
+        returns = []
+        for episode in range(cfg["EVAL"]["EPISODES"]):
+            state, done, ep_reward = env.reset(), False, 0.0
+            while not done:
+                action = agent.select_action(state)
+                state, reward, done, _ = env.step(action)
+                ep_reward += reward
+            returns.append(ep_reward)
+        print(f"Mean return over {len(returns)} episodes: {sum(returns)/len(returns)}")
+
+    elif args.mode == "demo":
+        agent.load(os.path.join(cfg["LOG"]["LOAD_DIR"], "agent.pth"))
+        state, done = env.reset(), False
+        while not done:
[…]

### 2025-04-20 traffic_signal_nn/utils/congif_parser.py

@@ -1,0 +1,29 @@
+# traffic_signal_nn/utils/config_parser.py
+
+import configparser
+
+def load_config(path):
+    """
+    Read INI file and return a dict-of-dicts: {SECTION: {KEY: value, …}, …}
+    Values are auto-cast to int/float/bool when possible.
+    """
+    parser = configparser.ConfigParser()
+    parser.read(path)
+    cfg = {}
+    for section in parser.sections():
+        d = {}
+        for key, val in parser[section].items():
+            d[key.upper()] = _auto_cast(val)
+        cfg[section.upper()] = d
+    return cfg
+
+def _auto_cast(val):
+    for caster in (int, float):
+        try:
+            return caster(val)
+        except ValueError:
+            pass
+    low = val.lower()
+    if low in ("true","false"):
+        return low == "true"
+    return val

### 2025-04-20 traffic_signal_nn/utils/logger.py

@@ -1,0 +1,25 @@
+# traffic_signal_nn/utils/logger.py
+
+import os
+from torch.utils.tensorboard import SummaryWriter
+
+class Logger:
+    """
+    Wrap TF‑Board and CSV logging.
+    """
+    def __init__(self, log_dir):
+        os.makedirs(log_dir, exist_ok=True)
+        self.writer = SummaryWriter(log_dir)
+        self.log_dir = log_dir
+
+    def log_episode(self, episode, reward, agent):
+        """
+        Log scalar metrics at end of each episode.
+        """
+        self.writer.add_scalar("Reward/episode", reward, episode)
+        # example: if agent tracks loss
+        if hasattr(agent, "loss"):
+            self.writer.add_scalar("Loss/td_loss", agent.loss, episode)
+
+    def close(self):
+        self.writer.close()

### 2025-04-20 traffic_signal_nn/utils/metrics.py

@@ -1,0 +1,18 @@
+# traffic_signal_nn/utils/metrics.py
+
+def compute_total_queue(env):
+    """
+    Sum of halting vehicles across all lanes.
+    """
+    return sum(env.traci.lane.getLastStepHaltingNumber(lane)
+               for lane in env.lanes)
+
+def compute_average_wait(env):
+    """
+    Average waiting time per vehicle.
+    """
+    vids = env.traci.vehicle.getIDList()
+    if not vids:
+        return 0.0
+    total = sum(env.traci.vehicle.getWaitingTime(v) for v in vids)
+    return total / len(vids)

### 2025-04-20 traffic_signal_nn/main.py
Fixed issues

@@ -1,9 +1,15 @@
+import sys, os
+#sys.path.insert(0, os.path.abspath(os.path.dirname(__file__)))
 import argparse
-import os
 from utils.config_parser import load_config
 from utils.logger import Logger
 from env.city_env import CityEnv
 from agents.policies import get_agent_class
+
+# from traffic_signal_nn.utils.config_parser import load_config
+# from traffic_signal_nn.utils.logger        import Logger
+# from traffic_signal_nn.env.city_env       import CityEnv
+# from traffic_signal_nn.agents.policies    import get_agent_class
 
 def main():
     parser = argparse.ArgumentParser(description="Multi‑RL Traffic Signal Control")

### 2025-04-21 paper/paper.tex
paper update

@@
 
-**INSERT IMAGE OF SUMO RUNNING**
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.9\linewidth]{SUMO_RUNNING.png}
+    \caption{Example of SUMO running in through a GUI}
+\end{figure}
 
@@
 
-Our first method of benchmark involves calculating the average waiting time of a vehicle. This is the most likely representation of how frustrated one might feel.
+Our first method of benchmark involves calculating the average waiting time of a vehicle.
+This is the most likely representation of how frustrated one might feel.
+A lower waiting time is therefore a great benchmark, as it shows that traffic is moving and drivers are less likely to get frustrated and develop road rage or get into an accident.
 
-The next is the average trip time
+The next is the average trip time.
+Becuase we will benchmark similar environments, we can get the average trip times of all combined vehicles in a simulation, and use it as a quantitative performance metric.
+Quicker trips indicates a successful traffic signal control implementation, as people can get to their end destination quicker, the entire point of driving.
+It also serves as a benchmark for other metrics, likely indicating shorter queue lengths, less stops, and less time waiting at stops on average.
 
-Finally we calculated information about queue length.
+Our final benchmark involves calculating queue lengths, or the average length of cars in line at an intersection.
+We can calulate both the max and average, to ensure the model is acting appropriately and not building super long lines to get the average down.
+A long line indicates that the model has a lot of buildup, and might fill later intersections with traffic, resulting in frustrated drivers and numerous stops if not carefully considered; therefore, it is critical that we minimize both the average queue length and the max queue length with our neural network.
+
+Notes on this...
 

### 2025-04-22 traffic_signal_nn/agents/models.py
Model Runs

@@ -1,14 +1,9 @@
-# traffic_signal_nn/agents/models.py
-
 import torch.nn as nn
 
-def build_mlp(input_dim, output_dim, hidden_units):
-    """
-    Construct a plain MLP: FC → ReLU → … → FC.
-    """
+def build_mlp(input_dim, output_dim, hidden_layers):
     layers = []
     prev = input_dim
-    for h in hidden_units:
+    for h in hidden_layers:
         layers.append(nn.Linear(prev, h))
         layers.append(nn.ReLU())
         prev = h

### 2025-04-22 traffic_signal_nn/agents/policies.py

@@ -1,85 +1,113 @@
 # traffic_signal_nn/agents/policies.py
-
 import random
 import torch
-import torch.optim as optim
 import torch.nn.functional as F
-from .models import build_mlp
-from .replay_buffer import ReplayBuffer
+import torch.optim as optim
 
-def get_agent_class(name):
-    name = name.lower()
-    if name == "dqn":
-        return DQNAgent
-    raise ValueError(f"Unknown agent '{name}'")
+from agents.models        import build_mlp
+from agents.replay_buffer import ReplayBuffer
 
+
+# ──────────────────────────────────────────────────────────────────────
 class DQNAgent:
-    """
-    Standard (single‑agent) DQN.
-    """
+    """One‑intersection DQN."""
     def __init__(self, state_dim, action_dim, cfg):
-        self.state_dim  = state_dim
-        self.action_dim = action_dim
-        self.lr         = cfg["LR"]
-        self.gamma      = cfg["GAMMA"]
-        self.epsilon    = cfg["EPS_START"]
-        self.eps_end    = cfg["EPS_END"]
-        self.eps_decay  = cfg["EPS_DECAY"]
+        hidden = list(map(int, cfg.get('HIDDEN', '64,64').split(',')))
+        self.net    = build_mlp(state_dim, action_dim, hidden)
+        self.target = build_mlp(state_dim, action_dim, hidden)
+        self.target.load_state_dict(self.net.state_dict())
 
-        hidden = list(map(int, cfg["HIDDEN"].split(",")))
-        self.policy_net = build_mlp(state_dim, action_dim, hidden)
-        self.target_net = build_mlp(state_dim, action_dim, hidden)
-        self.target_net.load_state_dict(self.policy_net.state_dict())
+        self.opt       = optim.Adam(self.net.parameters(),
+                                    lr=cfg.get('LR', 1e-3))
+        self.gamma     = cfg.get('GAMMA', 0.99)
+        self.eps       = cfg.get('EPS_START', 1.0)
+        self.eps_end   = cfg.get('EPS_END', 0.05)
+        self.eps_decay = cfg.get('EPS_DECAY', 0.995)
 
-        self.optimizer   = optim.Adam(self.policy_net.parameters(), lr=self.lr)
-        self.replay_buf  = ReplayBuffer(int(cfg["BUFFER_SIZE"]))
-        self.batch_size  = cfg["BATCH_SIZE"]
-        self.update_freq = cfg.get("TARGET_UPDATE_FREQ", 10)
-        self.step_count  = 0
[…]

### 2025-04-22 traffic_signal_nn/agents/replay_buffer.py

@@ -1,22 +1,16 @@
-# traffic_signal_nn/agents/replay_buffer.py
-
 import random
 from collections import deque
 
 class ReplayBuffer:
-    """
-    Fixed‑size FIFO buffer for experience replay.
-    """
     def __init__(self, capacity):
         self.buffer = deque(maxlen=capacity)
 
-    def add(self, state, action, reward, next_state, done):
-        self.buffer.append((state, action, reward, next_state, done))
+    def add(self, s, a, r, ns, done):
+        self.buffer.append((s, a, r, ns, done))
 
     def sample(self, batch_size):
         batch = random.sample(self.buffer, batch_size)
-        states, actions, rewards, next_states, dones = zip(*batch)
-        return list(states), list(actions), list(rewards), list(next_states), list(dones)
+        return zip(*batch)  # states, actions, rewards, next_states, dones
 
     def __len__(self):
         return len(self.buffer)

### 2025-04-22 traffic_signal_nn/env/city_env.py

@@ -1,32 +1,0 @@
-# traffic_signal_nn/env/city_env.py
-
-from .base_env import BaseEnv
-import traci
-
-class CityEnv(BaseEnv):
-    """
-    A single‑junction env for your target city.
-    """
-    def __init__(self, cfg):
-        """
-        cfg must include:
-          - SUMO_CFG: path to simulation.sumocfg
-          - TLS_ID:  traffic‑light system ID in the network
-          - N_PHASES: number of signal phases
-        """
-        super().__init__(cfg)
-        self.tls_id   = cfg["TLS_ID"]
-        self.n_phases = cfg["N_PHASES"]
-
-    def _apply_action(self, action):
-        # assume `action` is int phase index
-        traci.trafficlight.setPhase(self.tls_id, int(action))
-
-    @property
-    def observation_space(self):
-        # simple vector length = #lanes
-        return len(self.lanes)
-
-    @property
-    def action_space(self):
-        return self.n_phases

### 2025-04-22 traffic_signal_nn/env/multi_env.py

@@ -1,0 +1,43 @@
+import traci
+from .base_env import BaseEnv
+
+
+class CityEnv(BaseEnv):
+    """
+    Control *all* traffic lights (“fixed‑routes” scenario).
+    """
+
+    def __init__(self, cfg):
+        super().__init__(cfg)
+
+        raw = cfg.get('TLS_IDS', 'all').strip().lower()
+        self.tls_ids = (traci.trafficlight.getIDList()
+                        if raw == 'all'
+                        else [t.strip() for t in raw.split(',')])
+
+        self.n_phases = [len(traci.trafficlight
+                             .getAllProgramLogics(t)[0].phases)
+                         for t in self.tls_ids]
+
+    # ---- overrides ----
+    def _apply_action(self, actions):
+        if not isinstance(actions, (list, tuple)):
+            actions = [actions] * len(self.tls_ids)
+        for tid, act in zip(self.tls_ids, actions):
+            traci.trafficlight.setPhase(tid, int(act))
+
+    def get_state(self):
+        out = []
+        for tid in self.tls_ids:
+            lanes = traci.trafficlight.getControlledLanes(tid)
+            out.append([traci.lane.getLastStepHaltingNumber(l) for l in lanes])
+        return out
+
+    # ---- Gym spaces (simple) ----
+    @property
+    def observation_space(self):
+        return [len(x) for x in self.get_state()]
+
+    @property
+    def action_space(self):
+        return self.n_phases

### 2025-04-22 paper/paper.tex
Updated Paper

@@
 
-Notes on this...
+Because there are several benchmarks, we will consider our model successful if any evaluation metric improves without a meaningful reducing another metric.
+For example, if average trip time decreases, but the other metrics remain very similar to their threshold results, we will consider the model a success.
 

### 2025-04-22 traffic_signal_nn/agents/policies.py
NN complete

@@ -1,4 +1,3 @@
-# traffic_signal_nn/agents/policies.py
 import random
 import torch
 import torch.nn.functional as F
@@ -12,18 +11,25 @@ from agents.replay_buffer import ReplayBuffer
 class DQNAgent:
     """One‑intersection DQN."""
     def __init__(self, state_dim, action_dim, cfg):
+        # device setup
+        self.device = torch.device(
+            cfg.get('DEVICE', 'cuda' if torch.cuda.is_available() else 'cpu')
+        )
+        # network & target
         hidden = list(map(int, cfg.get('HIDDEN', '64,64').split(',')))
-        self.net    = build_mlp(state_dim, action_dim, hidden)
-        self.target = build_mlp(state_dim, action_dim, hidden)
+        self.net    = build_mlp(state_dim, action_dim, hidden).to(self.device)
+        self.target = build_mlp(state_dim, action_dim, hidden).to(self.device)
         self.target.load_state_dict(self.net.state_dict())
 
-        self.opt       = optim.Adam(self.net.parameters(),
-                                    lr=cfg.get('LR', 1e-3))
+        # optimizer & hyperparams
+        self.opt       = optim.Adam(self.net.parameters(), lr=cfg.get('LR', 1e-3))
         self.gamma     = cfg.get('GAMMA', 0.99)
         self.eps       = cfg.get('EPS_START', 1.0)
         self.eps_end   = cfg.get('EPS_END', 0.05)
         self.eps_decay = cfg.get('EPS_DECAY', 0.995)
+        self.action_dim = action_dim     # ① cache once
 
+        # replay buffer
         self.buf       = ReplayBuffer(cfg.get('BUFFER_SIZE', 10000))
         self.batch     = cfg.get('BATCH_SIZE', 32)
         self.update_fr = cfg.get('TARGET_UPDATE_FREQ', 10)
@@ -32,10 +38,11 @@ class DQNAgent:
     # --------------- public ----------------
     def select_action(self, state, greedy=False):
         if (not greedy) and random.random() < self.eps:
-            return random.randrange(self.net[-1].out_features)
-        with torch.no_grad():
-            q = self.net(torch.tensor(state, dtype=torch.float32))
-            return int(q.argmax())
+            return random.randrange(self.action_dim)          # ② cached value
+        with torch.inference_mode():                          # slightly faster
[…]

### 2025-04-22 traffic_signal_nn/env/base_env.py

@@ -1,49 +1,79 @@
 import traci
 
-
 class BaseEnv:
     """
-    Thin SUMO ↔ Gym‑like wrapper (fixed routes; one step = 1 s).
+    SUMO wrapper:
+        • sumo_dt  – internal SUMO step length   (0.1 s default)
+        • ctrl_dt  – RL control interval         (5 s  default)
     """
 
-    def __init__(self, cfg: dict):
-        self.sumo_bin = cfg.get('SUMO_BINARY', 'sumo')
-        self.sumocfg  = cfg['SUMO_CFG']
+    def __init__(self, cfg: dict, sumo_dt: float = 0.1, ctrl_dt: float = 5.0):
+        self.sumo_bin = cfg.get("SUMO_BINARY", "sumo")
+        self.sumocfg  = cfg["SUMO_CFG"]
+        self.sumo_dt  = float(sumo_dt)
+        self.ctrl_dt  = float(ctrl_dt)
 
-        self._start_cmd = [
-            self.sumo_bin, '-Q', '--no-step-log', '--no-warnings',
-            '--step-length', '1',              # ← here
-            '-c', self.sumocfg
+        COMMON = [
+            "--no-step-log",
+            "--no-warnings",
+            "--emergencydecel.warning-threshold", "9.99",
+            "--step-length",  str(self.sumo_dt),
+            "-c",             self.sumocfg,
         ]
-        self._load_cmd  = ['--step-length', '1', '-c', self.sumocfg]
+        traci.start([self.sumo_bin, "-Q", *COMMON])
+        self._load_cmd = COMMON[:]                         # used by traci.load()
 
-        traci.start(self._start_cmd)
-        self.lanes = traci.lane.getIDList()
+        # -------------- lane list PER traffic light -----------------
+        tls_ids = cfg.get("TLS_IDS", None)
+        if tls_ids and isinstance(tls_ids, str):
+            tls_ids = [tid.strip() for tid in tls_ids.split(",") if tid.strip()]
+        self.tls_ids  = tls_ids or [ ]                     # []  ⇒ single global agent
+        self.tls_lanes = (self._collect_tls_lanes()
+                          if self.tls_ids else [traci.lane.getIDList()])
 
-    # --------------- mandatory API ---------------
+    # ----------------------------------------------------------------
+    def _collect_tls_lanes(self):
+        """Return a list[ list[str] ]: lanes controlled by each TLS."""
+        all_lanes = traci.lane.getIDList()
[…]

### 2025-04-22 traffic_signal_nn/eval/evaluator.py

@@ -1,0 +1,85 @@
+import os
+import traci
+import sumolib
+import statistics
+import sys
+
+# Add the root directory (TrafficSignalManagement) to sys.path
+sys.path.append(os.path.abspath(os.path.join(os.path.dirname(__file__), '..')))
+
+from traffic_signal_nn.agents.policies import Agent
+from traffic_signal_nn.env.multi_env import CityEnv
+from traffic_signal_nn.utils.config_parser import load_config
+
+
+class Evaluator:
+    def __init__(self, cfg_path):
+        self.cfg = load_config(cfg_path)
+        self.env_cfg = self.cfg["ENV"]
+        self.eval_cfg = self.cfg["EVAL"]
+        self.agent_cfg = self.cfg["AGENT"]
+        self.model_dir = self.cfg["LOG"]["LOG_DIR"]
+
+        self.env = CityEnv(self.env_cfg)
+        self.agent = Agent(
+            self.env.observation_space,
+            self.env.action_space,
+            self.agent_cfg
+        )
+        self.load_models()
+
+    def load_models(self):
+        print("Loading trained models...")
+        for i, ag in enumerate(self.agent.agents):
+            path = os.path.join(self.model_dir, f"multi_dqn_tls{i}.pth")
+            if os.path.exists(path):
+                ag.load(path)
+            else:
+                print(f"⚠️ Missing model: {path}")
+
+    def evaluate(self):
+        print("Running evaluation...")
+        state, done = self.env.reset(), False
+
+        max_steps = 2400 # 1 hour of simulation time
+        steps = 0
+
+        while not done and steps < max_steps:
+            actions = self.agent.select_action(state, evaluate=True)
+            state, _, done, _ = self.env.step(actions)
+            steps += 2
+
+        # --- Extract SUMO Metrics ---
+        tripinfos = traci.simulation.getArrivedIDList()
+        durations = []
+        delays = []
+        waiting_times = []
+
+        for vid in tripinfos:
+            try:
+                durations.append(traci.vehicle.getAccumulatedWaitingTime(vid))
+                delays.append(traci.vehicle.getTimeLoss(vid))
+                waiting_times.append(traci.vehicle.getWaitingTime(vid))
+            except traci.exceptions.TraCIException:
+                continue
+
+        queue_lengths = [
+            traci.lane.getLastStepHaltingNumber(lane)
[…]

### 2025-04-22 traffic_signal_nn/main.py

@@ -30,37 +30,57 @@ def main() -> None:
     if args.mode == 'train':
         episodes = int(train_cfg.get('EPISODES', 0))
         for ep in range(episodes):
-            states, done, ep_ret = env.reset(), False, 0.0
-            while not done:
-                actions            = agent.select_action(states)
-                next_s, r, done, _ = env.step(actions)
+            ep_ret = 0.0
 
-                # broadcast r / done to per‑TLS lists
+            # ---------- episode starts ----------
+            print(f"[TRAIN] ▶ Episode {ep + 1}/{episodes} starting …", flush=True)
+            states, done = env.reset(), False 
+
+            max_steps = 2400 # 0.75 hour of simulation time
+            steps = 0
+
+            while not done and steps < max_steps:
+                actions            = agent.select_action(states)   # list[int]
+                next_s, rewards, done, _ = env.step(actions)       # rewards list[float]
+
                 n_tls = len(actions)
-                agent.remember(states, actions,
-                               [r] * n_tls, next_s,
-                               [done] * n_tls)
+                agent.remember(states, actions, rewards, next_s, [done]*n_tls)
                 agent.learn()
 
                 states = next_s
-                ep_ret += r
+                ep_ret += sum(rewards)      # aggregate episode return
+                steps  += 1
+
+            # ---------- episode ends ----------
             logger.log_episode(ep, ep_ret)
+            print(f"[TRAIN] ✓ Episode {ep + 1}/{episodes} finished — "
+                f"return: {ep_ret:.1f}", flush=True)
 
         agent.save(os.path.join(logger.log_dir,
                                 agent_cfg.get('NAME', 'dqn_city')))
-    else:                             # ---------- eval ----------
+
+    else:  # ------------------------ eval ------------------------
         episodes = int(eval_cfg.get('EPISODES', 0))
         returns  = []
         for ep in range(episodes):
+
+            print(f"[EVAL ] ▶ Episode {ep + 1}/{episodes} starting …", flush=True)
             s, done, tot = env.reset(), False, 0.0
+
             while not done:
                 a            = agent.select_action(s, evaluate=True)
                 s, r, done, _ = env.step(a)
                 tot         += r
+
             returns.append(tot)
-            print(f'Episode {ep}: {tot:.1f}')
+            print(f"[EVAL ] ✓ Episode {ep + 1}/{episodes} finished — "
[…]

### 2025-04-22 paper/paper.tex

@@
     \includegraphics[width=0.9\linewidth]{SUMO_RUNNING.png}
-    \caption{Example of SUMO running in through a GUI}
+    \caption{Example of SUMO running in through SUMO-GUI}
 \end{figure}
@@
 \subsection{Neural Network Implementation}
+
+our neural network is implemented through the python libraries traffic control interface (traci) and pytorch.
+Traci allows us to modify our simulation and get state information within a current timestep, while pytorch allows us to easily build and implement neural networks for our project.
+
+The first step in designing the proposed neural network is to establish what an agent is, and how they will communicate with the model.
+Based on existing codebases and research, we decided implementing a decentralized multi agent reinforcement learning model.
+This specific implementation lacks a central coordinator, and each agent acts on its own in coordination with other agents.
+The agents were designed to represent an intersection with a traffic signal controller, and could accept the state of the lanes leading to the intersection, the state of the signals, and the states of vehicles at each intersection such as stopped time; it also accepted a list of possible actions the traffic signal controller could perform, which when provided to a deep-q learning model can guess the optimal move.
+
+After designing an agent we decided to model the simulation environment in python to actually obtain states and actions for each agent.
+To do this, we used traci to obtain the states of vehicles on the lanes of traffic heading towards a specific traffic light, and assigned them to the traffic light.
+We then grabbed the possible state of actions, again assigning it to said traffic signal.
+This is neccessary for computing optimal actions later.
+We gather these every 5 seconds in the simulation, which greatly increases performance, and prevents rapidly switching signals during learning.
+
+Our intitial plan was to predict the optimal states of every intersection with a traffic signal controller in Starkville, MS, however, upon implementation we realized our available hardware and timeframe would not allow us to do this.
+Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly backed up.
+Most of these points were decided by considering the importance of the road, the time spent at intersections, and personal experience.
+This way we don't waste computational resources at intersections that see little traffic, or do fine directing traffic on their own.
+Doing so significantly decreases the computational requirements to run the model, and also allows us to focus only on important points.
+
+Our model does not have preset correct moves, which makes designing a neural network a very difficult task. Therefore, its important to choose a (non supervised), q-score, action, learn.
+
+implementing benchmarks
 

### 2025-04-23 traffic_signal_nn/agents/policies.py
Final Model

@@ -3,43 +3,37 @@ import torch
 import torch.nn.functional as F
 import torch.optim as optim
 
-from agents.models        import build_mlp
-from agents.replay_buffer import ReplayBuffer
+# Use relative imports within the traffic_signal_nn.agents package
+from .models import build_mlp
+from .replay_buffer import ReplayBuffer
 
-
-# ──────────────────────────────────────────────────────────────────────
 class DQNAgent:
-    """One‑intersection DQN."""
+    """One-intersection DQN."""
     def __init__(self, state_dim, action_dim, cfg):
-        # device setup
         self.device = torch.device(
             cfg.get('DEVICE', 'cuda' if torch.cuda.is_available() else 'cpu')
         )
-        # network & target
         hidden = list(map(int, cfg.get('HIDDEN', '64,64').split(',')))
         self.net    = build_mlp(state_dim, action_dim, hidden).to(self.device)
         self.target = build_mlp(state_dim, action_dim, hidden).to(self.device)
         self.target.load_state_dict(self.net.state_dict())
 
-        # optimizer & hyperparams
-        self.opt       = optim.Adam(self.net.parameters(), lr=cfg.get('LR', 1e-3))
-        self.gamma     = cfg.get('GAMMA', 0.99)
-        self.eps       = cfg.get('EPS_START', 1.0)
-        self.eps_end   = cfg.get('EPS_END', 0.05)
-        self.eps_decay = cfg.get('EPS_DECAY', 0.995)
-        self.action_dim = action_dim     # ① cache once
+        self.opt         = optim.Adam(self.net.parameters(), lr=cfg.get('LR', 1e-3))
+        self.gamma       = cfg.get('GAMMA', 0.99)
+        self.eps         = cfg.get('EPS_START', 1.0)
+        self.eps_end     = cfg.get('EPS_END', 0.05)
+        self.eps_decay   = cfg.get('EPS_DECAY', 0.995)
+        self.action_dim  = action_dim
 
-        # replay buffer
-        self.buf       = ReplayBuffer(cfg.get('BUFFER_SIZE', 10000))
-        self.batch     = cfg.get('BATCH_SIZE', 32)
-        self.update_fr = cfg.get('TARGET_UPDATE_FREQ', 10)
-        self.step_n    = 0
+        self.buf         = ReplayBuffer(cfg.get('BUFFER_SIZE', 10000))
[…]

### 2025-04-23 traffic_signal_nn/eval/evaluator.py

@@ -1,85 +1,0 @@
-import os
-import traci
-import sumolib
-import statistics
-import sys
-
-# Add the root directory (TrafficSignalManagement) to sys.path
-sys.path.append(os.path.abspath(os.path.join(os.path.dirname(__file__), '..')))
-
-from traffic_signal_nn.agents.policies import Agent
-from traffic_signal_nn.env.multi_env import CityEnv
-from traffic_signal_nn.utils.config_parser import load_config
-
-
-class Evaluator:
-    def __init__(self, cfg_path):
-        self.cfg = load_config(cfg_path)
-        self.env_cfg = self.cfg["ENV"]
-        self.eval_cfg = self.cfg["EVAL"]
-        self.agent_cfg = self.cfg["AGENT"]
-        self.model_dir = self.cfg["LOG"]["LOG_DIR"]
-
-        self.env = CityEnv(self.env_cfg)
-        self.agent = Agent(
-            self.env.observation_space,
-            self.env.action_space,
-            self.agent_cfg
-        )
-        self.load_models()
-
-    def load_models(self):
-        print("Loading trained models...")
-        for i, ag in enumerate(self.agent.agents):
-            path = os.path.join(self.model_dir, f"multi_dqn_tls{i}.pth")
-            if os.path.exists(path):
-                ag.load(path)
-            else:
-                print(f"⚠️ Missing model: {path}")
-
-    def evaluate(self):
-        print("Running evaluation...")
-        state, done = self.env.reset(), False
-
-        max_steps = 2400 # 1 hour of simulation time
-        steps = 0
-
-        while not done and steps < max_steps:
-            actions = self.agent.select_action(state, evaluate=True)
-            state, _, done, _ = self.env.step(actions)
-            steps += 2
-
-        # --- Extract SUMO Metrics ---
-        tripinfos = traci.simulation.getArrivedIDList()
-        durations = []
-        delays = []
-        waiting_times = []
-
-        for vid in tripinfos:
-            try:
-                durations.append(traci.vehicle.getAccumulatedWaitingTime(vid))
-                delays.append(traci.vehicle.getTimeLoss(vid))
-                waiting_times.append(traci.vehicle.getWaitingTime(vid))
-            except traci.exceptions.TraCIException:
-                continue
-
-        queue_lengths = [
-            traci.lane.getLastStepHaltingNumber(lane)
[…]

### 2025-04-23 traffic_signal_nn/eval/get_model_queue_length.py

@@ -1,0 +1,80 @@
+#!/usr/bin/env python3
+"""
+get_model_queue_length.py
+
+Runs a trained multi-DQN traffic signal controller in SUMO, collects the queue length
+(number of stopped vehicles at controlled lanes) at each control step, and writes the
+results to a CSV file.
+"""
+import os
+import sys
+import csv
+import argparse
+import statistics
+import traci
+
+# add project root so traffic_signal_nn package can be imported
+sys.path.insert(0, os.path.abspath(os.path.join(__file__, '..', '..', '..')))
+
+from traffic_signal_nn.env.multi_env      import CityEnv
+from traffic_signal_nn.utils.config_parser import load_config
+from traffic_signal_nn.agents.policies     import Agent
+
+
+def main():
+    p = argparse.ArgumentParser(description="Run model and log queue lengths per control step.")
+    p.add_argument('--config', required=True,
+                   help="path to traffic_signal_nn config .ini file")
+    p.add_argument('--output', default="./data/output/queue_length.csv",
+                   help="CSV file to write step,queue_length")
+    p.add_argument('--steps', type=int, default=None,
+                   help="maximum number of control steps to run")
+    args = p.parse_args()
+
+    # load config
+    cfg = load_config(args.config)
+    env_cfg = cfg['ENV']
+    agent_cfg = cfg.get('AGENT', {})
+    model_dir = cfg['LOG']['LOG_DIR']
+
+    # create environment and agent
+    env = CityEnv(env_cfg)
+    agent = Agent(env.observation_space, env.action_space, agent_cfg)
+
+    # load weights
+    print("Loading trained models …")
+    for i, ag in enumerate(agent.agents):
+        path = os.path.join(model_dir, f"multi_dqn_tls{i}.pth")
+        if os.path.exists(path):
+            ag.load(path)
+        else:
+            print(f"⚠️ Missing model: {path}")
+
+    # run simulation
+    print("Starting evaluation simulation …")
+    state, done = env.reset(), False
+    step = 0
+
+    with open(args.output, 'w', newline='') as csvfile:
+        writer = csv.writer(csvfile)
+        writer.writerow(["step", "queue_length"])
+
+        while not done and (args.steps is None or step < args.steps):
+            actions = agent.select_action(state, evaluate=True)
+            state, _, done, _ = env.step(actions)
+
+            # Compute queue-length across all controlled lanes:
+            # OPTION 2: use env.get_state()
[…]

### 2025-04-27 paper/paper.tex
paper updates

@@
 
-\title{Phase 2 - Literature Survey\\}
+\title{Phase 4 - Literature Survey and Results\\}
 
@@
 This tools allows us to run a visualization of the simulation with regards to the currently implemented method of traffic signal control, and make observations from the point of view of drivers without quantitative analysis, making observations about our model to make improvements.
+a
 
@@
 
-Our intitial plan was to predict the optimal states of every intersection with a traffic signal controller in Starkville, MS, however, upon implementation we realized our available hardware and timeframe would not allow us to do this.
+Our intitial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and timeframe would not allow us to do this.
 Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly backed up.
@@
 \section{Results over baseline}
+
+In order to evaluate our model we did two tests, a qualitative analysis, which involves viewing the simulation to compare the original simulation to the baseline, and quantitative analysis where benchmarks were ran to evaluate the model with numbers representing the time taken driving and length of queues.
+Both methods are critical to effective analysis of the model.
+While training involved generating various traffic scenarios, or events where cars start at one point and end at another point, our testing involves one scenario, which uses similar traffic patterns seen to training.
+Using a single traffic scenario allows us to make observations of threshold and neural network based signals.
+This ensures that while testing, each vehicle has takes the same route so we can observe how our traffic signal management system performs relative to other methods.
+
+\subsection{qualitative analysis}
+
+When observing the baseline threshold based traffic signals, it did an acceptable job managing light-medium traffic (2000 vehicles per hour), but saw relative redundance with lights changing while no cars were in the opposite direction, causing congestion on major roads.
+It felt familiar, with minor traffic on less congested roads, such as backroads connected to highways and highways without traffic stops, but was certainly not as effective as possible.
+
+The MARL Deep Q-learning model saw noteable improvements in terms of responsiveness.
+When there was no signs of incoming traffic or lines in stoped lanes, it skipped the redunancy of changing the state of traffic lights resulting in a generally smoother flow of traffic.
+When it did change, it seemed intelligent, considering how long cars have been waiting along with how many cars were waiting.
+While our model did not allow collisions to be considered, the more continuous flow of traffic saw less emergency stops, and generally safer roads.
+Analysis also revealed that vehicles would often arrive at their destination sooner than their threshold counterparts, with less time spent on the road generally resulting in safer roads with less congestion.
+
+There were some limitations of MARL Deep Q-learning despite its generally better analysis.
+Oftentimes, a light would not change for long periods of times to direct more intense traffic at the expense of shorter lines.
+This would generally result in shorter average drive lengths, with one drive taking exceptionally long.
+This was relatively controlled, however, with no cars getting completely stuck, and eventually being directed even if it took a minute.
+It was also important to choose the right intersections as points of interest, as intersections with little to no congestion would flip their signals the moment a car came to an intersection.
+It was generally better for areas with extremely light traffic to stick with threshold signals.
+For points like these, a potentially better option is sensor based traffic signal controllers, which use more concrete rules to change traffic, but make observations similar to our learning model to inform these decisions.
+
+\subsection{quantitative analysis}
+
+While simple observations provide great insight to the performance of our model, numbers to support our observations are potentially more important.
+As mentioned earlier, we developed several benchmarks based on the results of a simulation.
+Each benchmark provides a real insight that indicates the perforamnce of a traffic signal control method without the need to make observations.
+
+The first benchmark was..
+
+the seconds
+
+fourth
+
+last
+
+\subsection{Analysis of overall results}
+
+yadieaisdfj;asdkf
 

### 2025-04-27 paper/paper.tex
Update paper.tex
basic copyedit

@@
 While a powerful tool, it struggled to handle the data format provided from OSM data, resulting in slow performance and broken intersections.
-We then decided moving to the SUMO simulator was a more worthwhile decision, as its efficency has improved over time and documentation is more appropriate for our specific research area.
+We then decided moving to the SUMO simulator was a more worthwhile decision, as its efficiency has improved over time and documentation is more appropriate for our specific research area.
 
-Once SUMO was installed, we downloaded public data for the city of Starkville Mississippi's road structure using the tool open street map, and used SUMO's netedit to process the data into a useable format.
+Once SUMO was installed, we downloaded public data for the city of Starkville Mississippi's road structure using the tool open street map, and used SUMO's netedit to process the data into a usable format.
 Some changes were made, such as fixing traffic signal placement using junctions, and fixing non-signal based intersection interactions.
@@
 
-In order to properly evaluate the performance of threshold and neural network models, a method of evaluation was needed beyond qualitative methods, such as observation or reivew.
+In order to properly evaluate the performance of threshold and neural network models, a method of evaluation was needed beyond qualitative methods, such as observation or review.
 As a result we implemented various benchmarks to quantitatively decide the quality of a methods approach.
@@
 The next is the average trip time.
-Becuase we will benchmark similar environments, we can get the average trip times of all combined vehicles in a simulation, and use it as a quantitative performance metric.
+Because we will benchmark similar environments, we can get the average trip times of all combined vehicles in a simulation, and use it as a quantitative performance metric.
 Quicker trips indicates a successful traffic signal control implementation, as people can get to their end destination quicker, the entire point of driving.
@@
 Our final benchmark involves calculating queue lengths, or the average length of cars in line at an intersection.
-We can calulate both the max and average, to ensure the model is acting appropriately and not building super long lines to get the average down.
+We can calculate both the max and average, to ensure the model is acting appropriately and not building super long lines to get the average down.
 A long line indicates that the model has a lot of buildup, and might fill later intersections with traffic, resulting in frustrated drivers and numerous stops if not carefully considered; therefore, it is critical that we minimize both the average queue length and the max queue length with our neural network.
@@
 We then grabbed the possible state of actions, again assigning it to said traffic signal.
-This is neccessary for computing optimal actions later.
+This is necessary for computing optimal actions later.
 We gather these every 5 seconds in the simulation, which greatly increases performance, and prevents rapidly switching signals during learning.
 
-Our intitial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and timeframe would not allow us to do this.
+Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and timeframe would not allow us to do this.
 Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly backed up.
@@
 
-\subsection{qualitative analysis}
+\subsection{Qualitative Analysis}
 
-When observing the baseline threshold based traffic signals, it did an acceptable job managing light-medium traffic (2000 vehicles per hour), but saw relative redundance with lights changing while no cars were in the opposite direction, causing congestion on major roads.
+When observing the baseline threshold-based traffic signals, it did an acceptable job managing light-medium traffic (2000 vehicles per hour), but saw relative redundancy with lights changing while no cars were in the opposite direction, causing congestion on major roads.
 It felt familiar, with minor traffic on less congested roads, such as backroads connected to highways and highways without traffic stops, but was certainly not as effective as possible.
 
-The MARL Deep Q-learning model saw noteable improvements in terms of responsiveness.
-When there was no signs of incoming traffic or lines in stoped lanes, it skipped the redunancy of changing the state of traffic lights resulting in a generally smoother flow of traffic.
+The MARL Deep Q-learning model saw notable improvements in terms of responsiveness.
+When there was no signs of incoming traffic or lines in stopped lanes, it skipped the redundancy of changing the state of traffic lights resulting in a generally smoother flow of traffic.
 When it did change, it seemed intelligent, considering how long cars have been waiting along with how many cars were waiting.
@@
 
-\subsection{quantitative analysis}
+\subsection{Quantitative Analysis}
 
@@
 As mentioned earlier, we developed several benchmarks based on the results of a simulation.
-Each benchmark provides a real insight that indicates the perforamnce of a traffic signal control method without the need to make observations.
+Each benchmark provides a real insight that indicates the performance of a traffic signal control method without the need to make observations.
 
@@
 
-\subsection{Analysis of overall results}
+\subsection{Analysis of Overall Results}
 

### 2025-04-27 paper/paper.tex
Update paper.tex

@@
 Although Du et al.
-have begun to account for pedestrians \cite{13}, there are still several factors that go unaccounted for in the process of traffic signal prediction, which can back up traffic worse than threshold options.
+have begun to take into account pedestrians \cite{13}, there are still several factors that remain unaccounted for in the traffic signal prediction process, which can back up traffic worse than threshold options.
 
@@
 
-After beginning implementation we decided that CityFlow was not the appropriate tool for this problem, despite our research.
+After starting implementation, we decided that CityFlow was not the appropriate tool for this problem, despite our research.
 While a powerful tool, it struggled to handle the data format provided from OSM data, resulting in slow performance and broken intersections.
@@
 Once SUMO was installed, we downloaded public data for the city of Starkville Mississippi's road structure using the tool open street map, and used SUMO's netedit to process the data into a usable format.
-Some changes were made, such as fixing traffic signal placement using junctions, and fixing non-signal based intersection interactions.
+Some changes were made, such as fixing traffic signal placement using junctions and fixing non-signal-based intersection interactions.
 
@@
 
-Using the setup configuration file, we can use SUMO GUI, which comes bundled with the installation package of SUMO.
+Using the configuration file, we can use the SUMO GUI, which comes included with the SUMO installation package.
 This tools allows us to run a visualization of the simulation with regards to the currently implemented method of traffic signal control, and make observations from the point of view of drivers without quantitative analysis, making observations about our model to make improvements.
@@
 Because we will benchmark similar environments, we can get the average trip times of all combined vehicles in a simulation, and use it as a quantitative performance metric.
-Quicker trips indicates a successful traffic signal control implementation, as people can get to their end destination quicker, the entire point of driving.
+Faster trips indicate a successful implementation of traffic signal control, as people can get to their end destination faster.
 It also serves as a benchmark for other metrics, likely indicating shorter queue lengths, less stops, and less time waiting at stops on average.
@@
 
-our neural network is implemented through the python libraries traffic control interface (traci) and pytorch.
+Our neural network is implemented through the python libraries traffic control interface (traci) and pytorch.
 Traci allows us to modify our simulation and get state information within a current timestep, while pytorch allows us to easily build and implement neural networks for our project.
@@
 The first step in designing the proposed neural network is to establish what an agent is, and how they will communicate with the model.
-Based on existing codebases and research, we decided implementing a decentralized multi agent reinforcement learning model.
+Based on existing codebases and research, we decided to implement a decentralized multi-agent reinforcement learning model.
 This specific implementation lacks a central coordinator, and each agent acts on its own in coordination with other agents.
@@
 
-Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and timeframe would not allow us to do this.
+Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and time frame would not allow us to do this.
 Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly backed up.

### 2025-04-27 paper/paper.tex
Update paper.tex

@@
 
-Our neural network is implemented through the python libraries traffic control interface (traci) and pytorch.
-Traci allows us to modify our simulation and get state information within a current timestep, while pytorch allows us to easily build and implement neural networks for our project.
+Our decentralized multi‐agent reinforcement learning (MARL) architecture is built around three key components: the Q‐network itself, a DQN‐based agent wrapper (complete with replay buffer and learning logic), and a SUMO–traci environment interface that supplies state and reward information every control interval.
+Traci allows us to easily modify our simulation and get state information from within the current timestep, while pytorch allows us to efficiently build and implement our neural networks.
+This implementation lacks a central coordinator.
+Each agent acts on its own in coordination with other agents.
+The agents were designed to represent an intersection with a traffic signal controller, and could accept the state of the lanes leading to the intersection, the state of the signals, and the states of vehicles at each intersection such as stopped time.
+We define the neural network in models.py as a simple multilayer perceptron.
+Given an input dimension equal to the total number of lanes feeding into a signalized intersection and an output dimension equal to the number of possible signal phases, we stack two hidden layers of configurable width (48 units each in our experiments) with ReLU activations, and terminate with a linear layer producing a Q‐value for each phase.
 
-The first step in designing the proposed neural network is to establish what an agent is, and how they will communicate with the model.
-Based on existing codebases and research, we decided to implement a decentralized multi-agent reinforcement learning model.
-This specific implementation lacks a central coordinator, and each agent acts on its own in coordination with other agents.
-The agents were designed to represent an intersection with a traffic signal controller, and could accept the state of the lanes leading to the intersection, the state of the signals, and the states of vehicles at each intersection such as stopped time; it also accepted a list of possible actions the traffic signal controller could perform, which when provided to a deep-q learning model can guess the optimal move.
+Concretely, the helper function
+\begin{verbatim}
+build_mlp(input_dim, output_dim, hidden_layers):
+    layers   = []
+    prev     = input_dim
+    for h in hidden_layers:
+        layers.append(nn.Linear(prev, h))
+        layers.append(nn.ReLU())
+        prev = h
+    layers.append(nn.Linear(prev, output_dim))
+    return nn.Sequential(*layers)
+\end{verbatim}
+constructs the sequential network, allowing us to easily swap hidden‐layer sizes or depths via configuration.
 
-After designing an agent we decided to model the simulation environment in python to actually obtain states and actions for each agent.
-To do this, we used traci to obtain the states of vehicles on the lanes of traffic heading towards a specific traffic light, and assigned them to the traffic light.
-We then grabbed the possible state of actions, again assigning it to said traffic signal.
-This is necessary for computing optimal actions later.
-We gather these every 5 seconds in the simulation, which greatly increases performance, and prevents rapidly switching signals during learning.
+Each intersection is managed by an instance of {\tt DQNAgent} (in policies.py), which maintains both an \emph{online} network and a \emph{target} network.
+At initialization, we load the target’s weights from the online network, and set up an Adam optimizer (learning rate 0.001), a discount factor $\gamma=0.99$, and an $\varepsilon$‐greedy exploration schedule that decays $\varepsilon$ from 1.0 to 0.05 by a factor of 0.995 per learning step.
+Experiences are stored in a fixed‐capacity deque buffer (size 5,000), and we sample mini‐batches of 32 transitions once the buffer is sufficiently large.
+In each learn() call we compute the predicted Q‐values for state–action pairs, form TD‐targets via \(Q_{\mathrm{target}} = r + \gamma (1 - d)\max_{a'} Q_{\mathrm{target}}(s', a')\) and minimize the mean‐squared error between predictions and targets.
+After each gradient update we decay $\varepsilon$, and every 100 updates we synchronize the target network with the online network.
+To scale this to multiple intersections, we wrap several {\tt DQNAgent} objects in {\tt MultiDQNAgent}, forwarding action selection, memory storage, and learning calls independently to each agent.
 
-Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and time frame would not allow us to do this.
-Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly backed up.
-Most of these points were decided by considering the importance of the road, the time spent at intersections, and personal experience.
-This way we don't waste computational resources at intersections that see little traffic, or do fine directing traffic on their own.
-Doing so significantly decreases the computational requirements to run the model, and also allows us to focus only on important points.
+On the environment side (multi\_env.py, subclassing BaseEnv), we employ the python library traci (traffic control interface) to simulate traffic in Starkville, MS.
+Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and the time frame would not allow us to do this.
+Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly heavy and prone to congestion.
+Most of these points were decided by considering the importance of the road, the typical time spent at intersections, and personal experience.
+This method ensures that we don't expend computational resources at intersections that either see little traffic or do fine managing traffic on their own.
+Doing so significantly decreases the computational requirements to run the model, and allows us to focus only on critical points.
+At startup we invoke traci.start with a step length of 0.1s and a control interval of 5s, disabling verbose logs to save overhead.
+We automatically collect, for each traffic‐light signal (TLS), the set of incoming lanes by inspecting traci.trafficlight.getControlledLinks.
+Every control step, get\_state() returns a list of queue lengths, i.e.\ the number of halting vehicles per lane, by calling traci.lane.getLastStepHaltingNumber on each lane.
+We then compute a scalar reward for each TLS as the negative sum of its queue lengths, which incentivizes the agents to minimize congestion.
+When step(actions) is called, we apply each agent’s chosen phase via traci.trafficlight.setPhase, advance the simulation by the control interval, and then return the new state, reward list, and a done flag (triggered when no vehicles remain).
 
-Our model does not have preset correct moves, which makes designing a neural network a very difficult task. Therefore, its important to choose a (non supervised), q-score, action, learn.
-
-implementing benchmarks
+The overall training loop lives in main.py, which begins by parsing our configuration file.
+This configuration specifies 500 training episodes, an agent name ``multi\_dqn,'' and all key hyperparameters (learning rate, $\gamma$, $\varepsilon$ schedule, hidden‐layer sizes, buffer and batch sizes, and target‐update frequency).
+For each episode, we reset the environment, then for up to 2,400 steps (equivalent to 0.75 hours of simulated time) we query the multi‐agent for actions, step the environment, store transitions, and trigger a learning update immediately.
+Episode returns (the sum of all TLS rewards) are logged each episode, and upon completion of training we save each agent’s weights to disk.
+A similar loop runs in evaluation mode (with greedy action selection and no learning) to report average performance over ten episodes.
 

### 2025-04-27 paper/paper.tex
Update paper.tex

@@
 
-Our decentralized multi‐agent reinforcement learning (MARL) architecture is built around three key components: the Q‐network itself, a DQN‐based agent wrapper (complete with replay buffer and learning logic), and a SUMO–traci environment interface that supplies state and reward information every control interval.
-Traci allows us to easily modify our simulation and get state information from within the current timestep, while pytorch allows us to efficiently build and implement our neural networks.
+Our decentralized multi‐agent reinforcement learning (MARL) architecture is built around three key components: the Q‐network itself, a Deep Q-Network(DQN)‐based agent wrapper (complete with replay buffer and learning logic), and a SUMO–TraCI environment interface that supplies state and reward information every control interval.
+Traci allows us to easily modify our simulation and get state information from within the current timestep, while PyTorch allows us to efficiently build and implement our neural networks.
 This implementation lacks a central coordinator.
@@
 
-On the environment side (multi\_env.py, subclassing BaseEnv), we employ the python library traci (traffic control interface) to simulate traffic in Starkville, MS.
+On the environment side (multi\_env.py, subclassing BaseEnv), we employ the Python library TraCI (traffic control interface) to simulate traffic in Starkville, MS.
 Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and the time frame would not allow us to do this.

### 2025-04-27 paper/paper.tex
Update paper.tex

@@
 Research on traffic signal methods falls into three categories.
-These include single agent learning models, multi agent reinforcement learning models (MARL), and large-scale multi-agent learning models, with the large-scale MARL actively having new deep learning models applied in recent research.
+These include single-agent learning models, multi-agent reinforcement learning models (MARL), and large-scale multi-agent learning models, with the large-scale MARL actively having new deep learning models applied in recent research.
 
-\subsection{Single Agent Learning Models}
+\subsection{Single-agent Learning Models}
 
-Before 2010, traffic prediction models were designed using single agent reinforcement learning models.
+Before 2010, traffic prediction models were designed using single-agent reinforcement learning models.
 These models were used to predict the movement of traffic using a single system, meaning a single neural network predicted how traffic lights should function.
@@
 
-\subsection{Multi Agent Learning Models}
+\subsection{Multi-agent Learning Models}
 
@@
 
-During this research period, the majority of models used a variation of Q-Learning.
+During this research period, the majority of models used a variation of Q-learning.
 While not perfect, there were indications that there were improvements in terms of moving traffic over traditional timer methods.
@@
 
-Since 2019, research indicates that MARL agents have been improved for scalability, performance, real-world applications and become open-source.
-The field has seen enormous growth from the initial small scale solutions, and large scale implementations for real world cities continue to become more practical and powerful.
+Since 2019, research indicates that MARL agents have been improved for scalability, performance, and real-world applications and become open-source.
+The field has seen enormous growth from the initial small-scale solutions, and large-scale implementations for real-world cities continue to become more practical and powerful.
 
 Early breakthroughs in large-scale MARL focused on a more difficult problem: urban traffic management.
-This initially came with a new open sourced model, CityFlow, a multi-agent reinforcement model which could handle large road networks with thousands of intersections in a simulated environment \cite{7}.
+This initially came with a new open-sourced model, CityFlow, a multi-agent reinforcement model which could handle large road networks with thousands of intersections in a simulated environment \cite{7}.
 These large-scale models have shown significant improvements in performance while working at a larger scale, making them more adaptable to real situations.
@@
 In recent years, strong reinforcement learning models, such as long short-term memory (LSTM) and spatiotemporal graph convolutional neural networks (STGCN) have been applied to research, and external factors like pedestrian movements have begun being addressed.
-These advancements show more promise in scalability and performance, but are still in progress.
+These advancements show more promise in scalability and performance but are still in progress.
 There are still opportunities to test different models which may perform better in synchronization with other agents.
-While accounting for pedestrians marks a major step in optimizing the traffic signal problem, it opens more questions, such as how can we account for emergency vehicles, namely ambulances or police cars, or collisions that disturb the natural flow of traffic.
+While accounting for pedestrians marks a major step in optimizing the traffic signal problem, it opens more questions, such as how we can account for emergency vehicles, namely ambulances or police cars, or collisions that disturb the natural flow of traffic.
 Despite these questions, these models are very powerful in comparison to previous threshold traffic signal management systems.
@@
 Based on the major research contributions, we have concluded that the papers fall into several categories.
-These categories tend to reflect the time period of the research contribution, with older papers using less-powerful Q-Learning simulations, and more recent papers using more powerful neural networks with account for more factors.
+These categories tend to reflect the time period of the research contribution, with older papers using less powerful Q-learning simulations, and more recent papers using more powerful neural networks that account for more factors.
 Figure 2 highlights a chronological overview of the categorization of papers over time.
@@
 
-Another research gap is regards real-world application.
+Another research gap regards real-world application.
 Several large cities have adapted deep learning techniques for traffic signal processing, though a large majority of cities do not yet have these technologies implemented.
@@
 However, there are a variety of challenges with implementation.
-One challenge is the difficulty in finding optimal Q values as there is no ground truth to compare it to, the agent has to guess and learn on the experience and rewards it receives.
+One challenge is the difficulty in finding optimal Q values as there is no ground truth to compare it to, the agent has to guess and learn from the experience and rewards it receives.
 This method is also very computationally challenging, as the agent has to constantly update the Q-value, especially with very large and complex models.
@@
 There are also problems that could arise with the environment.
-The environment can slowly change overtime, so our Q-value may become irrelevant, so the agent would have to relearn its environment.
-There are also many unexpected situations that may arise that may require the model to quickly learn and adapt in the unknown situation, or try to reinforce the model to account for these situations.
+The environment can slowly change over time, so our Q-value may become irrelevant, so the agent would have to relearn its environment.
+There are also many unexpected situations that may arise that may require the model to quickly learn and adapt to the unknown situation, or try to reinforce the model to account for these situations.
 
@@
 After starting implementation, we decided that CityFlow was not the appropriate tool for this problem, despite our research.
-While a powerful tool, it struggled to handle the data format provided from OSM data, resulting in slow performance and broken intersections.
+While a powerful tool, it struggled to handle the data format provided by OSM data, resulting in slow performance and broken intersections.
 We then decided moving to the SUMO simulator was a more worthwhile decision, as its efficiency has improved over time and documentation is more appropriate for our specific research area.
 
-Once SUMO was installed, we downloaded public data for the city of Starkville Mississippi's road structure using the tool open street map, and used SUMO's netedit to process the data into a usable format.
+Once SUMO was installed, we downloaded public data for the city of Starkville Mississippi's road structure using the tool OpenStreetMap and used SUMO's netedit to process the data into a usable format.
 Some changes were made, such as fixing traffic signal placement using junctions and fixing non-signal-based intersection interactions.
 
-Our ultimate goal with this project is to test how neural network influenced traffic signals direct the flow of traffic, so we needed to setup various simulation environments to simulate realistic traffic scenarios within Starkville.
-To do this, we used SUMO's built in flow generator, with various parameters to result in light, medium, and intense traffic scenarios.
-This would allow us to test how each method of traffic signal control reacts to various events such as morning and afternoon rush hours, post game traffic, or light summer traffic.
+Our ultimate goal with this project is to test how neural network-influenced traffic signals direct the flow of traffic, so we needed to set up various simulation environments to simulate realistic traffic scenarios within Starkville.
+To do this, we used SUMO's built-in flow generator, with various parameters to result in light, medium, and intense traffic scenarios.
+This would allow us to test how each method of traffic signal control reacts to various events such as morning and afternoon rush hours, post-game traffic, or light summer traffic.
 
 Using the configuration file, we can use the SUMO GUI, which comes included with the SUMO installation package.
-This tools allows us to run a visualization of the simulation with regards to the currently implemented method of traffic signal control, and make observations from the point of view of drivers without quantitative analysis, making observations about our model to make improvements.
+This tool allows us to run a visualization of the simulation with regards to the currently implemented method of traffic signal control, and make observations from the point of view of drivers without quantitative analysis, making observations about our model to make improvements.
 a
@@
 In order to properly evaluate the performance of threshold and neural network models, a method of evaluation was needed beyond qualitative methods, such as observation or review.
-As a result we implemented various benchmarks to quantitatively decide the quality of a methods approach.
+As a result, we implemented various benchmarks to quantitatively decide the quality of a method's approach.
 
-Our first method of benchmark involves calculating the average waiting time of a vehicle.
+Our first method of benchmarking involves calculating the average waiting time of a vehicle.
 This is the most likely representation of how frustrated one might feel.
@@
 Faster trips indicate a successful implementation of traffic signal control, as people can get to their end destination faster.
-It also serves as a benchmark for other metrics, likely indicating shorter queue lengths, less stops, and less time waiting at stops on average.
+It also serves as a benchmark for other metrics, likely indicating shorter queue lengths, fewer stops, and less time waiting at stops on average.
 
@@
 
-Because there are several benchmarks, we will consider our model successful if any evaluation metric improves without a meaningful reducing another metric.
-For example, if average trip time decreases, but the other metrics remain very similar to their threshold results, we will consider the model a success.
+Because there are several benchmarks, we will consider our model successful if any evaluation metric improves without a meaningful reduction of another metric.
+For example, if the average trip time decreases, but the other metrics remain very similar to their threshold results, we will consider the model a success.
 
@@
 Each agent acts on its own in coordination with other agents.
-The agents were designed to represent an intersection with a traffic signal controller, and could accept the state of the lanes leading to the intersection, the state of the signals, and the states of vehicles at each intersection such as stopped time.
+The agents were designed to represent an intersection with a traffic signal controller and could accept the state of the lanes leading to the intersection, the state of the signals, and the state of vehicles at each intersection such as stopped time.
 We define the neural network in models.py as a simple multilayer perceptron.
-Given an input dimension equal to the total number of lanes feeding into a signalized intersection and an output dimension equal to the number of possible signal phases, we stack two hidden layers of configurable width (48 units each in our experiments) with ReLU activations, and terminate with a linear layer producing a Q‐value for each phase.
+Given an input dimension equal to the total number of lanes feeding into a signalized intersection and an output dimension equal to the number of possible signal phases, we stack two hidden layers of configurable width (48 units each in our experiments) with ReLU activations and terminate with a linear layer producing a Q‐value for each phase.
 
@@
 \end{verbatim}
-constructs the sequential network, allowing us to easily swap hidden‐layer sizes or depths via configuration.
+constructs the sequential network, allowing us to easily swap hidden layer sizes or depths via configuration.
 
 Each intersection is managed by an instance of {\tt DQNAgent} (in policies.py), which maintains both an \emph{online} network and a \emph{target} network.
-At initialization, we load the target’s weights from the online network, and set up an Adam optimizer (learning rate 0.001), a discount factor $\gamma=0.99$, and an $\varepsilon$‐greedy exploration schedule that decays $\varepsilon$ from 1.0 to 0.05 by a factor of 0.995 per learning step.
+At initialization, we load the target’s weights from the online network and set up an Adam optimizer (learning rate 0.001), a discount factor $\gamma=0.99$, and an $\varepsilon$‐greedy exploration schedule that decays $\varepsilon$ from 1.0 to 0.05 by a factor of 0.995 per learning step.
 Experiences are stored in a fixed‐capacity deque buffer (size 5,000), and we sample mini‐batches of 32 transitions once the buffer is sufficiently large.
 In each learn() call we compute the predicted Q‐values for state–action pairs, form TD‐targets via \(Q_{\mathrm{target}} = r + \gamma (1 - d)\max_{a'} Q_{\mathrm{target}}(s', a')\) and minimize the mean‐squared error between predictions and targets.
-After each gradient update we decay $\varepsilon$, and every 100 updates we synchronize the target network with the online network.
+After each gradient update, we decay $\varepsilon$, and every 100 updates we synchronize the target network with the online network.
 To scale this to multiple intersections, we wrap several {\tt DQNAgent} objects in {\tt MultiDQNAgent}, forwarding action selection, memory storage, and learning calls independently to each agent.
@@
 On the environment side (multi\_env.py, subclassing BaseEnv), we employ the Python library TraCI (traffic control interface) to simulate traffic in Starkville, MS.
-Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and the time frame would not allow us to do this.
+Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation, we realized our available hardware and the time frame would not allow us to do this.
 Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly heavy and prone to congestion.
@@
 We automatically collect, for each traffic‐light signal (TLS), the set of incoming lanes by inspecting traci.trafficlight.getControlledLinks.
-Every control step, get\_state() returns a list of queue lengths, i.e.\ the number of halting vehicles per lane, by calling traci.lane.getLastStepHaltingNumber on each lane.
+For every control step, get\_state() returns a list of queue lengths, i.e.\ the number of halting vehicles per lane, by calling traci.lane.getLastStepHaltingNumber on each lane.
 We then compute a scalar reward for each TLS as the negative sum of its queue lengths, which incentivizes the agents to minimize congestion.
@@
 For each episode, we reset the environment, then for up to 2,400 steps (equivalent to 0.75 hours of simulated time) we query the multi‐agent for actions, step the environment, store transitions, and trigger a learning update immediately.
-Episode returns (the sum of all TLS rewards) are logged each episode, and upon completion of training we save each agent’s weights to disk.
+Episode returns (the sum of all TLS rewards) are logged each episode, and upon completion of training, we save each agent’s weights to disk.
 A similar loop runs in evaluation mode (with greedy action selection and no learning) to report average performance over ten episodes.
@@
 
-In order to evaluate our model we did two tests, a qualitative analysis, which involves viewing the simulation to compare the original simulation to the baseline, and quantitative analysis where benchmarks were ran to evaluate the model with numbers representing the time taken driving and length of queues.
-Both methods are critical to effective analysis of the model.
-While training involved generating various traffic scenarios, or events where cars start at one point and end at another point, our testing involves one scenario, which uses similar traffic patterns seen to training.
-Using a single traffic scenario allows us to make observations of threshold and neural network based signals.
-This ensures that while testing, each vehicle has takes the same route so we can observe how our traffic signal management system performs relative to other methods.
+In order to evaluate our model we did two tests, a qualitative analysis, which involves viewing the simulation to compare the original simulation to the baseline, and a quantitative analysis where benchmarks were run to evaluate the model with numbers representing the time taken driving and length of queues.
+Both methods are critical to the effective analysis of the model.
+While training involves generating various traffic scenarios, or events where cars start at one point and end at another point, our testing involves one scenario, which uses similar traffic patterns seen in training.
+Using a single traffic scenario allows us to make observations of threshold and neural network-based signals.
+This ensures that while testing, each vehicle takes the same route so we can observe how our traffic signal management system performs relative to other methods.
 
@@
 The MARL Deep Q-learning model saw notable improvements in terms of responsiveness.
-When there was no signs of incoming traffic or lines in stopped lanes, it skipped the redundancy of changing the state of traffic lights resulting in a generally smoother flow of traffic.
-When it did change, it seemed intelligent, considering how long cars have been waiting along with how many cars were waiting.
-While our model did not allow collisions to be considered, the more continuous flow of traffic saw less emergency stops, and generally safer roads.
+When there were no signs of incoming traffic or lines in stopped lanes, it skipped the redundancy of changing the state of traffic lights resulting in a generally smoother flow of traffic.
+When it did change, it seemed intelligent, considering how long cars had been waiting along with how many cars were waiting.
+While our model did not allow collisions to be considered, the more continuous flow of traffic saw fewer emergency stops, and generally safer roads.
 Analysis also revealed that vehicles would often arrive at their destination sooner than their threshold counterparts, with less time spent on the road generally resulting in safer roads with less congestion.
@@
 There were some limitations of MARL Deep Q-learning despite its generally better analysis.
-Oftentimes, a light would not change for long periods of times to direct more intense traffic at the expense of shorter lines.
+Oftentimes, a light would not change for long periods of time to direct more intense traffic at the expense of shorter lines.
 This would generally result in shorter average drive lengths, with one drive taking exceptionally long.
@@
 It was generally better for areas with extremely light traffic to stick with threshold signals.
-For points like these, a potentially better option is sensor based traffic signal controllers, which use more concrete rules to change traffic, but make observations similar to our learning model to inform these decisions.
+For points like these, a potentially better option is sensor-based traffic signal controllers, which use more concrete rules to change traffic but make observations similar to our learning model to inform these decisions.
 
@@
 
-While simple observations provide great insight to the performance of our model, numbers to support our observations are potentially more important.
+While simple observations provide great insight into the performance of our model, numbers to support our observations are potentially more important.
 As mentioned earlier, we developed several benchmarks based on the results of a simulation.
@@
 
-The other sources provide great insight, but might have had a smaller research impact.
+The other sources provide great insight but might have had a smaller research impact.
 Despite this, they are still high-quality papers and can be used to help solve our problem.
-These are still critical to understand the problem at hand, and have a meaningful impact on development of our traffic prediction model.
+These are still critical to understanding the problem at hand and have a meaningful impact on the development of our traffic prediction model.
 

### 2025-04-27 paper/paper.tex
Update paper.tex

@@
     \includegraphics[width=0.9\linewidth]{SUMO_RUNNING.png}
-    \caption{Example of SUMO running in through SUMO-GUI}
+    \caption{Example of SUMO running in SUMO-GUI}
 \end{figure}
@@
 
+\newpage 
 Concretely, the helper function
 \begin{verbatim}
-build_mlp(input_dim, output_dim, hidden_layers):
+build_mlp(input_dim, output_dim, 
+hidden_layers):
     layers   = []
@@
         prev = h
-    layers.append(nn.Linear(prev, output_dim))
+    layers.append(nn.Linear(prev, 
+    output_dim))
     return nn.Sequential(*layers)
@@
 Experiences are stored in a fixed‐capacity deque buffer (size 5,000), and we sample mini‐batches of 32 transitions once the buffer is sufficiently large.
-In each learn() call we compute the predicted Q‐values for state–action pairs, form TD‐targets via \(Q_{\mathrm{target}} = r + \gamma (1 - d)\max_{a'} Q_{\mathrm{target}}(s', a')\) and minimize the mean‐squared error between predictions and targets.
+In each learn() call we compute the predicted Q‐values for state–action pairs, form TD‐targets via
+\[
+Q_{\mathrm{target}} \;=\; r \;+\; \gamma\,(1 - d)\,\max_{a'} Q_{\mathrm{target}}(s', a')
+\]
+and minimize the mean‐squared error between predictions and targets.
 After each gradient update, we decay $\varepsilon$, and every 100 updates we synchronize the target network with the online network.

### 2025-04-28 paper/paper.tex
paper progress

@@
 \usepackage{xcolor}
-\usepackage{float} % Add this in your preamble if not included
+\usepackage{float} 
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 have begun to take into account pedestrians \cite{13}, there are still several factors that remain unaccounted for in the traffic signal prediction process, which can back up traffic worse than threshold options.
-
-\section*{Work Distribution}
-The majority of research findings were made by Jason Weeks and Andrew McBride with help from Andrei by branching out from the original CityFlow paper \cite{7} and finding predecessors and related articles in the field.
-
-Based on the research findings, Jason Weeks, Jacquies Turner, and Andrew McBride contributed to the research taxonomy and timelines.
-These were made by narrowing down the most important research articles in our original findings.
-
-This paper was primarily written by Jason Weeks, but all the members contributed.
-
-The final distribution of the work is as follows.
-
-\begin{itemize}
-    \item 25\% - Jason Weeks
-    \item 25\% - Andrew McBride
-    \item 25\% - Jacquies Turner
-    \item 25\% - Andrei Roskelley Garcia
-\end{itemize}
 
@@
 
-The first benchmark was..
+Testing the average trip time was one of the most important benchmarks implemented, since longer trips result in less cars on the road and a shorter time spent on the road, the odds of a traffic incident decreases dramatically.
+When using traditional threshold signals the average trip time took 695.34 seconds, or 11 minutes and 34 seconds.
+When using neural network based traffic signals a noteable decrease in average trip time was observed, decreasing to 587.61 seconds, or 9 minutes and 47 seconds.
+With a nearly 108 second decrease in average trip time, we saw a 15\% reduction in average trip time.
 
-the seconds
+We also analyzed time spent waiting for both benchmarks, to see where the decrease in average trip time came from.
+In theory, this result would come from a decreased wait time, but some of it could come from the decreased traffic and increased speeds associated with better traffic signal management.
+With threshold signals, the average time spent waiting was 298.4 seconds, but with neural networks this decreased to 195.76 seconds.
+This resulted in a decrease of 102 seconds, confirming that the decreased wait time significantly contributed to the lower average trip time.
 
-fourth
+Another important metric to decrease it the average length of stopped cars at any point in time, known as the average queue length.
+Because our wait time decreased, we can expect a noteable decrease in average queue length, as less cars need to stop and wait.
+This is confirmed, as we achieve an average queue length of 230 vehicles with threshold neural networks, and a queue length of 55 vehicles when using neural networks.
+This is an astronomical decrease, with an average of 175 less vehicles being stopped at any point in time.
+This is a nearly 76\% decrease in average queue length, indicating a significant imporvement in performance over baselines.
 
-last
+Fianlly, the maximum length of cars was determined in order to ensure that traffic did not ever back up very bad.
+Because the average queue length decreased greatly, it is safe to assume that the maximum queue length decreased significantly, which is confirmed by our results.
+When using traditional threshold signals the maximum queue length reached 1373 vehicles, and the maximum queue length reached 316 vehicles.
+This indicates that the peak congestion was significantly improved through our model over baselines, preventing large backups which can be frustrating and potentially dangerous.
+
+Each of the benchmarks and their comparison to the baseline can be seen below.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|l|c|c|c|}
+\hline
+\textbf{Metric} & \textbf{Without Q-Learning} & \textbf{MARL Deep Q-Learning} & \textbf{Change} \\
+\hline
+Avg. Trip Time & 695.34 seconds & 587.61 seconds & ↓107.33 \\
+\hline
+Avg. Waiting Time & 298.40 seconds & 195.76 seconds & ↓102.64 \\
+\hline
+Avg. Queue Length & 230 vehicles & 54.98 vehicles & ↓175.02 \\
+\hline
+Max Queue Length & 1373 vehicles & 316 vehicles & ↓1057 \\
+\hline
+\end{tabular}
+\caption{Performance comparison between Without Q-Learning and MARL Deep Q-Learning.}
+\label{tab:comparison}
+\end{table}
 
@@
 
-yadieaisdfj;asdkf
+Both qualitative and quantitative analysis reveal that deep Q learning is effective at modeling the traffic signal controllers within Starkville Mississippi, and for both traffic speeds and safety the Mississippi Department of Transport (MDOT) should consider updating the traffic signal control methods for points where traffic gets particularly congested.
+
+With the average time spent driving decreasing drivers are generally safer, and large sections of traffic are less likely to make sudden stops.
+The shorter queue lengths result in less frustrated drivers, and quicker driving times, which also indicates a smoother flow of traffic and safer roads.
+This did come at the cost of computational power, resulting in a slower model that took time to train.
+It does perform faster than real time, but given the wrong situation could slow down the model where it could not make real time predictions; however, because the results were overall positive, there likely exists a cheaper and more efficient way to direct traffic.
+
+While deep Q-Learning with MARL objectively performed well, computation times were relatively slow and implementation would prove very costly.
+Because of this, the MDOT should also investigate sensor based traffic signal controllers.
+These signals are cheaper to install, but consider timers and congestion levels with more strictly defined rules.
+Implementing these would avoid the high cost of training, implementation, and upkeep.
+Because using state data increased the performance with neural networks, its likely that a simple understanding of the state of the road could help direct traffic without implementing an expensive and unpredictable model.
+While still having some cost, it would decrease time spent driving and keep the roads of Mississippi safer.
 
@@
 Not every paper analyzed was used.
+
+\section*{Work Distribution}
+The majority of research findings were made by Jason Weeks and Andrew McBride with help from Andrei by branching out from the original CityFlow paper \cite{7} and finding predecessors and related articles in the field.
+
+Based on the research findings, Jason Weeks, Jacquies Turner, and Andrew McBride contributed to the research taxonomy and timelines.
+These were made by narrowing down the most important research articles in our original findings.
+
+The final distribution of the work is as follows.
+
+\begin{itemize}
+    \item 25\% - Jason Weeks
+    \item 25\% - Andrew McBride
+    \item 25\% - Jacquies Turner
+    \item 25\% - Andrei Roskelley Garcia
+\end{itemize}
 

### 2025-04-29 paper/paper.tex

@@
 
-\newpage 
 Concretely, the helper function
@@
 When using traditional threshold signals the average trip time took 695.34 seconds, or 11 minutes and 34 seconds.
-When using neural network based traffic signals a noteable decrease in average trip time was observed, decreasing to 587.61 seconds, or 9 minutes and 47 seconds.
+When using neural network based traffic signals a notable decrease in average trip time was observed, decreasing to 587.61 seconds, or 9 minutes and 47 seconds.
 With a nearly 108 second decrease in average trip time, we saw a 15\% reduction in average trip time.
@@
 Another important metric to decrease it the average length of stopped cars at any point in time, known as the average queue length.
-Because our wait time decreased, we can expect a noteable decrease in average queue length, as less cars need to stop and wait.
+Because our wait time decreased, we can expect a notable decrease in average queue length, as less cars need to stop and wait.
 This is confirmed, as we achieve an average queue length of 230 vehicles with threshold neural networks, and a queue length of 55 vehicles when using neural networks.
 This is an astronomical decrease, with an average of 175 less vehicles being stopped at any point in time.
-This is a nearly 76\% decrease in average queue length, indicating a significant imporvement in performance over baselines.
+This is a nearly 76\% decrease in average queue length, indicating a significant improvement in performance over baselines.
 
-Fianlly, the maximum length of cars was determined in order to ensure that traffic did not ever back up very bad.
+Finally, the maximum length of cars was determined in order to ensure that traffic did not ever back up very bad.
 Because the average queue length decreased greatly, it is safe to assume that the maximum queue length decreased significantly, which is confirmed by our results.

### 2025-05-06 paper/paper.tex
paper progress

@@
 The overall training loop lives in main.py, which begins by parsing our configuration file.
-This configuration specifies 500 training episodes, an agent name ``multi\_dqn,'' and all key hyperparameters (learning rate, $\gamma$, $\varepsilon$ schedule, hidden‐layer sizes, buffer and batch sizes, and target‐update frequency).
+This configuration specifies 200 training episodes, an agent name ``multi\_dqn,'' and all key hyperparameters (learning rate, $\gamma$, $\varepsilon$ schedule, hidden‐layer sizes, buffer and batch sizes, and target‐update frequency).
 For each episode, we reset the environment, then for up to 2,400 steps (equivalent to 0.75 hours of simulated time) we query the multi‐agent for actions, step the environment, store transitions, and trigger a learning update immediately.
@@
 A similar loop runs in evaluation mode (with greedy action selection and no learning) to report average performance over ten episodes.
+
+\subsection{Cross Validation}
+
+One potential concern over the implementation of our model is not the model itself, but the trustworthiness of the results.
+As a result, the design of the model took careful consideration into preparing valid results through cross validation and critical analysis.
+Various situations were considered for training, while only a single traffic scenario was used for testing and benhcmarking.
+
+The first way we validated our results is by creating new scenarios.
+Each Episode of the model sees a completely new traffic scenario within Starkville, MS, granted with similar size due to computational limitations of large scale traffic.
+We then printed the reward, the punishment part of the Q-score each episode the ensure progress and exploration, and the results reveal relative consistency with a strong balance of exploration and exploitation.
+Even with new situations, the score grew and grew over time (please note that the reward is negative).
+The model was trained for 150 episodes in this scenario, to reach convergance while exploring as much as possible.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{Reward.png}
+    \caption{Reward over 150 Episodes}
+\end{figure}
+
+During our benchmarking phase we generated one testing scenario.
+This scenario uses a moderate-light traffic, based on Starkville, Mississippi's population and is used as a baseline to ensure that the improved results are not a result of a more or less congested traffic pattern.
+When we run the benchmarks we carefully ensured that this was performed to make sure results are trustworthy, and the the model did not opimize only one single situation it saw.
+When doing this, the results were overwhelmingly positive, and ensures that MARL based Deep Q-Learning is a valid way to optimize traffic signal controllers, as seen in section ``Results over baseline.``
 
@@
 
-\begin{table}[h!]
+\begin{table}[H]
 \centering
@@
 \hline
-\textbf{Metric} & \textbf{Without Q-Learning} & \textbf{MARL Deep Q-Learning} & \textbf{Change} \\
+\textbf{Metric} & \textbf{Without Q-Learning} & \textbf{Deep Q-Learning} & \textbf{Change} \\
 \hline
-Avg. Trip Time & 695.34 seconds & 587.61 seconds & ↓107.33 \\
+Trip Time & 695.34 seconds & 587.61 seconds & ↓107.33 \\
 \hline
-Avg. Waiting Time & 298.40 seconds & 195.76 seconds & ↓102.64 \\
+Waiting Time & 298.40 seconds & 195.76 seconds & ↓102.64 \\
 \hline
-Avg. Queue Length & 230 vehicles & 54.98 vehicles & ↓175.02 \\
+Avg. Queue& 230 vehicles & 54.98 vehicles & ↓175.02 \\
 \hline
-Max Queue Length & 1373 vehicles & 316 vehicles & ↓1057 \\
+Max Queue & 1373 vehicles & 316 vehicles & ↓1057 \\
 \hline
 \end{tabular}
-\caption{Performance comparison between Without Q-Learning and MARL Deep Q-Learning.}
+\caption{Performance comparison between Without Q-Learning and Deep Q-Learning.}
 \label{tab:comparison}
@@
 
+\section{The Codebase and Demoing}
+
+\subsection{Installation}
+
+In order to download the code one would need to go to github, and view chandlerweeks/TrafficSignalManagement.
+The link to the repository is: https://github.com/ChandlerWeeks/TrafficSignalManagement.
+Once here, fork the repository and make a local clone on your device.
+If a release is available at the time of reading the paper, that link will allow direct access of the codebase.
+
+The model was not designed within a virtual environment, due to conflicts between operating systems of installations, so one will need to install the dependencies for the code beforehand, most noteably traci, likely via the pip install traci through a terminal.
+
+In order to run the code, SUMO will also need to be installed, following SUMO's tutorial for installation for the necessary operating system.
+Once installed, the path variables will need to be set for sumo and sumo-gui in order for the code to run.
+
+\subsection{Live Demo}
+
+Once every prerequisite has been setup, to test installation and configuration, one should open SUMO-gui and open the file in data/simulation.sumocfg.
+If the file runs and traffic is observed, then the testing traffic flow scenario has been loaded and the code is ready to run.
+
+-- running training
+
+-- running testing
+
 \section{ACKNOWLEDGMENT OF LLM USE}
@@
 Not every paper analyzed was used.
+
+Large language models were used in assistance of coding, particularly for debugging and assistance with small code snippets, but never used to generate large parts of a funciton, or codebits.
+The code is developed by human, but AI assisted with some lines.
 

### 2025-05-07 paper/paper.tex
more paper work

@@
 
--- running training
+In order to run training, navigate to the TrafficSignalManagement file in a terminal instance and run the provided command.
+\texttt{\texttt{python traffic\_signal\_nn/main.py --config traffic\_signal\_nn/config/config\_dqn\_city.ini train}}.
+Common errors include not having SUMO variables added to the path, and not having pip modules installed.
+Each episode will run, taking about a minute to simulate an hour of traffic.
+The reward will be printed to the console to let the user observe the loss, to see when the model explores and when it converges.
+The code can take hours at a time to run, and a demo can be provided if necessary, via youtube.
 
--- running testing
+In order to benchmark the baseline for SUMO, the scripts benchmark.py and get\_queue\_length.py were developed.
+benchmark simply gets the results of the runtime through the file data/output/tripinfo.xml.
+It contains information about the trip, such as average trip time, average waiting time, and average delay, but does not contain information on average queue length.
+Before running benchmarks, it is important to run the get queue length script, which obtains the average queue lengths seperate from what SUMO provides.
+One must run the benchmark scenario provided through simulation.sumocfg within SUMO-gui to get the correct output.
 

### 2025-05-07 traffic_signal_nn/env/base_env.py
small code revision

@@ -1,12 +1,6 @@
 import traci
 
 class BaseEnv:
-    """
-    SUMO wrapper:
-        • sumo_dt  – internal SUMO step length   (0.1 s default)
-        • ctrl_dt  – RL control interval         (5 s  default)
-    """
-
     def __init__(self, cfg: dict, sumo_dt: float = 0.1, ctrl_dt: float = 5.0):
         self.sumo_bin = cfg.get("SUMO_BINARY", "sumo")
         self.sumocfg  = cfg["SUMO_CFG"]

### 2025-05-07 traffic_signal_nn/eval/run_model.py

@@ -1,16 +1,8 @@
 #!/usr/bin/env python3
-"""
-run_model.py
-
-Load a trained multi-DQN traffic signal agent and run the SUMO simulation until completion.
-"""
 import os
 import sys
 import argparse
 import traci
-
-# Make sure project root is on PYTHONPATH when running as module
-#sys.path.append(os.path.abspath(os.path.join(__file__, '..')))
 
 from traffic_signal_nn.env.multi_env      import CityEnv
 from traffic_signal_nn.utils.config_parser import load_config

### 2025-05-07 paper/paper.tex

@@
 The reward will be printed to the console to let the user observe the loss, to see when the model explores and when it converges.
+Parameters can be modified for training within traffic\_signal\_nn/config/config\_dqn\_city.ini, including the neural network parameters and the selected roads.
 The code can take hours at a time to run, and a demo can be provided if necessary, via youtube.
@@
 One must run the benchmark scenario provided through simulation.sumocfg within SUMO-gui to get the correct output.
+In order to run the benchmarks, simply run \texttt{python benchmark.py} after running the simulation, along with \texttt{python .\textbackslash get\_queue\_length.py --config .\textbackslash data\textbackslash simulation.sumocfg --output .\textbackslash data\textbackslash output\textbackslash queue\_length.csv} to determine the queue lengths.
+
+In order to run the benchmarks with the model, one must run the trained model first to get the trip information.
+This can be ran through this command \texttt{python -m traffic\_signal\_nn.eval.run\_model --config traffic\_signal\_nn/config/config\_dqn\_city.ini}, which runs the trained model on the same situation as the previous benchmark, for consistency as mentioned in cross-validation.
+Just like without the neural network, the queue length is not obtained for the benchmarks, so one will need to run \texttt{python .\textbackslash traffic\_signal\_nn\textbackslash eval\textbackslash get\_model\_queue\_length.py --config .\textbackslash traffic\_signal\_nn\textbackslash config\textbackslash config\_dqn\_city.ini --output .\textbackslash data\textbackslash output\textbackslash queue\_length.csv} in order the get the updated queue lengths.
+Once all of this has been done, one can simply run \texttt{python benchmark.py} to get the benchmark results for the trained model.
+
+Setup of the project can be rather time consuming, as mentioned, a video can provided to the reviewers running the script completely setup if necessary.
 

### 2025-05-07 paper/paper.tex
Update paper.tex

@@
 and minimize the mean‐squared error between predictions and targets.
+This target arises directly from the Bellman optimality equation and serves as the reference value toward which our online Q‐network is trained.
+Below is a concise, technical explanation of each component.
+The scalar \textit{r} represents the instantaneous reward signal returned by the environment upon executing action \textit{a} in state \textit{s}.
+In traffic‐signal control, we typically define \textit{r} as a negative congestion metric (e.g., the sum of queue lengths), so that higher rewards correspond to smoother traffic flow.
+This term anchors the TD target in the actual, observed outcome of the agent’s decision.
+$\gamma \in [0,1)$ is the temporal‐discount coefficient that geometrically attenuates the contribution of future returns.
+A value of $\gamma = 0.99$ indicates that the agent values rewards received one step later at $99\%$ of their original magnitude.  
+By tuning $\gamma$, one controls the trade‐off between short‐term gains and long‐horizon optimization.
+Lower $\gamma$ leads to myopic behavior, while values approaching 1 emphasize long‐term planning.
+The binary flag $d$ equals $1$ if the sampled transition terminates an episode (i.e., $s'$ is terminal), and $0$ otherwise.
+Multiplying by $(1 - d)$ ensures that, in terminal states, no bootstrapped future value is added, thus strictly enforcing that $Q_{\mathrm{target}} = r$ when the episode ends.
+This prevents spurious propagation of post‐terminal $Q$-values.
+The operator $\max_{a'} Q(s',a')$ extracts the highest $Q$‐value estimate across all possible actions $a'$ in the successor state $s'$, as evaluated by the target network.
+This “greedy” selection embodies the assumption that, after the current transition, the agent will henceforth behave optimally, consistently choosing the action with maximal expected return.
+By summing the immediate reward $r$ and the discounted maximal next‐state value (when $d = 0$), we obtain a single‐step bootstrap estimate of the true return.
+During training, the online network’s prediction $Q_{\mathrm{pred}}(s,a)$ is regressed toward $Q_{\mathrm{target}}$ via mean‐squared error.
+Periodic synchronization of the target network stabilizes training by holding the bootstrap reference fixed for a configurable number of gradient updates.
 After each gradient update, we decay $\varepsilon$, and every 100 updates we synchronize the target network with the online network.

### 2025-05-07 paper/paper.tex
Update paper.tex

@@
 As a result, the design of the model took careful consideration into preparing valid results through cross validation and critical analysis.
-Various situations were considered for training, while only a single traffic scenario was used for testing and benhcmarking.
+Various situations were considered for training, while only a single traffic scenario was used for testing and benchmarking.
 
@@
 Even with new situations, the score grew and grew over time (please note that the reward is negative).
-The model was trained for 150 episodes in this scenario, to reach convergance while exploring as much as possible.
+The model was trained for 150 episodes in this scenario, to reach convergence while exploring as much as possible.
 
@@
 This scenario uses a moderate-light traffic, based on Starkville, Mississippi's population and is used as a baseline to ensure that the improved results are not a result of a more or less congested traffic pattern.
-When we run the benchmarks we carefully ensured that this was performed to make sure results are trustworthy, and the the model did not opimize only one single situation it saw.
+When we run the benchmarks we carefully ensured that this was performed to make sure results are trustworthy, and the the model did not optimize only one single situation it saw.
 When doing this, the results were overwhelmingly positive, and ensures that MARL based Deep Q-Learning is a valid way to optimize traffic signal controllers, as seen in section ``Results over baseline.``
@@
 
-In order to download the code one would need to go to github, and view chandlerweeks/TrafficSignalManagement.
+In order to download the code one would need to go to GitHub, and view chandlerweeks/TrafficSignalManagement.
 The link to the repository is: https://github.com/ChandlerWeeks/TrafficSignalManagement.
@@
 
-The model was not designed within a virtual environment, due to conflicts between operating systems of installations, so one will need to install the dependencies for the code beforehand, most noteably traci, likely via the pip install traci through a terminal.
+The model was not designed within a virtual environment, due to conflicts between operating systems of installations, so one will need to install the dependencies for the code beforehand, most notably traci, likely via the pip install traci through a terminal.
 
 In order to run the code, SUMO will also need to be installed, following SUMO's tutorial for installation for the necessary operating system.
-Once installed, the path variables will need to be set for sumo and sumo-gui in order for the code to run.
+Once installed, the path variables will need to be set for sumo and sumo-GUI in order for the code to run.
 
@@
 
-Once every prerequisite has been setup, to test installation and configuration, one should open SUMO-gui and open the file in data/simulation.sumocfg.
+Once every prerequisite has been setup, to test installation and configuration, one should open SUMO-GUI and open the file in data/simulation.sumocfg.
 If the file runs and traffic is observed, then the testing traffic flow scenario has been loaded and the code is ready to run.
@@
 Parameters can be modified for training within traffic\_signal\_nn/config/config\_dqn\_city.ini, including the neural network parameters and the selected roads.
-The code can take hours at a time to run, and a demo can be provided if necessary, via youtube.
+The code can take hours at a time to run, and a demo can be provided if necessary, via YouTube.
 
@@
 It contains information about the trip, such as average trip time, average waiting time, and average delay, but does not contain information on average queue length.
-Before running benchmarks, it is important to run the get queue length script, which obtains the average queue lengths seperate from what SUMO provides.
-One must run the benchmark scenario provided through simulation.sumocfg within SUMO-gui to get the correct output.
+Before running benchmarks, it is important to run the get queue length script, which obtains the average queue lengths separate from what SUMO provides.
+One must run the benchmark scenario provided through simulation.sumocfg within SUMO-GUI to get the correct output.
 In order to run the benchmarks, simply run \texttt{python benchmark.py} after running the simulation, along with \texttt{python .\textbackslash get\_queue\_length.py --config .\textbackslash data\textbackslash simulation.sumocfg --output .\textbackslash data\textbackslash output\textbackslash queue\_length.csv} to determine the queue lengths.
@@
 
-Large language models were used in assistance of coding, particularly for debugging and assistance with small code snippets, but never used to generate large parts of a funciton, or codebits.
+Large language models were used in assistance of coding, particularly for debugging and assistance with small code snippets, but never used to generate large parts of a function, or codebits.
 The code is developed by human, but AI assisted with some lines.

### 2025-05-07 paper/paper.tex
Update paper.tex

@@
 \section{Introduction}
-This paper will present the relevant literature in the process of traffic signal control using machine learning techniques, along with the timeline of research developments over time.
+This paper will present the relevant literature on the process of traffic signal control using machine learning techniques, along with the timeline of research developments over time.
 
@@
 Early breakthroughs in large-scale MARL focused on a more difficult problem: urban traffic management.
-This initially came with a new open-sourced model, CityFlow, a multi-agent reinforcement model which could handle large road networks with thousands of intersections in a simulated environment \cite{7}.
+This initially came with a new open-sourced model, CityFlow, a multi-agent reinforcement model that could handle large road networks with thousands of intersections in a simulated environment \cite{7}.
 These large-scale models have shown significant improvements in performance while working at a larger scale, making them more adaptable to real situations.
@@
 
-A third research gap appears to be the influence of various real-world distractions that can occur, ranging from road accidents, pedestrians, and emergency vehicles.
+A third research gap appears to be the influence of various real-world distractions that can occur, ranging from road accidents, to pedestrians, and emergency vehicles.
 Although Du et al.
@@
 The operator $\max_{a'} Q(s',a')$ extracts the highest $Q$‐value estimate across all possible actions $a'$ in the successor state $s'$, as evaluated by the target network.
-This “greedy” selection embodies the assumption that, after the current transition, the agent will henceforth behave optimally, consistently choosing the action with maximal expected return.
+This “greedy” selection embodies the assumption that, after the current transition, the agent will henceforth behave optimally, consistently choosing the action with the maximal expected return.
 By summing the immediate reward $r$ and the discounted maximal next‐state value (when $d = 0$), we obtain a single‐step bootstrap estimate of the true return.
@@
 
-\subsection{Cross Validation}
+\subsection{Cross-Validation}
 
 One potential concern over the implementation of our model is not the model itself, but the trustworthiness of the results.
-As a result, the design of the model took careful consideration into preparing valid results through cross validation and critical analysis.
+As a result, the design of the model took careful consideration into preparing valid results through cross-validation and critical analysis.
 Various situations were considered for training, while only a single traffic scenario was used for testing and benchmarking.
@@
 The first way we validated our results is by creating new scenarios.
-Each Episode of the model sees a completely new traffic scenario within Starkville, MS, granted with similar size due to computational limitations of large scale traffic.
-We then printed the reward, the punishment part of the Q-score each episode the ensure progress and exploration, and the results reveal relative consistency with a strong balance of exploration and exploitation.
+Each episode of the model sees a completely new traffic scenario within Starkville, MS, granted of a similar size due to computational limitations of large-scale traffic simulation.
+We then printed the reward, and the punishment part of the Q-score for each episode to ensure progress and exploration, and the results reveal relative consistency with a strong balance of exploration and exploitation.
 Even with new situations, the score grew and grew over time (please note that the reward is negative).
@@
 
-During our benchmarking phase we generated one testing scenario.
-This scenario uses a moderate-light traffic, based on Starkville, Mississippi's population and is used as a baseline to ensure that the improved results are not a result of a more or less congested traffic pattern.
-When we run the benchmarks we carefully ensured that this was performed to make sure results are trustworthy, and the the model did not optimize only one single situation it saw.
-When doing this, the results were overwhelmingly positive, and ensures that MARL based Deep Q-Learning is a valid way to optimize traffic signal controllers, as seen in section ``Results over baseline.``
+During our benchmarking phase, we generated one testing scenario.
+This scenario uses moderate-light traffic, based on Starkville, Mississippi's population, and is used as a baseline to ensure that the improved results are not a result of a more or less congested traffic pattern.
+When we ran the benchmarks we carefully ensured that this was performed to make sure results were trustworthy, and the model did not optimize only one single situation it saw.
+When doing this, the results were overwhelmingly positive, and ensures that MARL-based Deep Q-Learning is a valid way to optimize traffic signal controllers, as seen in the section ``Results over baseline.``
 
@@
 While our model did not allow collisions to be considered, the more continuous flow of traffic saw fewer emergency stops, and generally safer roads.
-Analysis also revealed that vehicles would often arrive at their destination sooner than their threshold counterparts, with less time spent on the road generally resulting in safer roads with less congestion.
+The analysis also revealed that vehicles would often arrive at their destination sooner than their threshold counterparts, with less time spent on the road generally resulting in safer roads with less congestion.
 
@@
 
-Testing the average trip time was one of the most important benchmarks implemented, since longer trips result in less cars on the road and a shorter time spent on the road, the odds of a traffic incident decreases dramatically.
+Testing the average trip time was one of the most important benchmarks implemented, since longer trips resulted in fewer cars on the road and a shorter time spent on the road, the odds of a traffic incident decreased dramatically.
 When using traditional threshold signals the average trip time took 695.34 seconds, or 11 minutes and 34 seconds.
-When using neural network based traffic signals a notable decrease in average trip time was observed, decreasing to 587.61 seconds, or 9 minutes and 47 seconds.
-With a nearly 108 second decrease in average trip time, we saw a 15\% reduction in average trip time.
+When using neural network-based traffic signals a notable decrease in average trip time was observed, decreasing to 587.61 seconds, or 9 minutes and 47 seconds.
+With a nearly 108-second decrease in average trip time, we saw a 15\% reduction in average trip time.
 
@@
 In theory, this result would come from a decreased wait time, but some of it could come from the decreased traffic and increased speeds associated with better traffic signal management.
-With threshold signals, the average time spent waiting was 298.4 seconds, but with neural networks this decreased to 195.76 seconds.
+With threshold signals, the average time spent waiting was 298.4 seconds, but with neural networks, this decreased to 195.76 seconds.
 This resulted in a decrease of 102 seconds, confirming that the decreased wait time significantly contributed to the lower average trip time.
 
-Another important metric to decrease it the average length of stopped cars at any point in time, known as the average queue length.
-Because our wait time decreased, we can expect a notable decrease in average queue length, as less cars need to stop and wait.
+Another important metric to decrease is the average length of stopped cars at any point in time, known as the average queue length.
+Because our wait time decreased, we can expect a notable decrease in average queue length, as fewer cars need to stop and wait.
 This is confirmed, as we achieve an average queue length of 230 vehicles with threshold neural networks, and a queue length of 55 vehicles when using neural networks.
-This is an astronomical decrease, with an average of 175 less vehicles being stopped at any point in time.
+This is an astronomical decrease, with an average of 175 fewer vehicles being stopped at any point in time.
 This is a nearly 76\% decrease in average queue length, indicating a significant improvement in performance over baselines.
 
-Finally, the maximum length of cars was determined in order to ensure that traffic did not ever back up very bad.
+Finally, the maximum length of cars was determined in order to ensure that traffic did not ever back up very badly.
 Because the average queue length decreased greatly, it is safe to assume that the maximum queue length decreased significantly, which is confirmed by our results.
@@
 This did come at the cost of computational power, resulting in a slower model that took time to train.
-It does perform faster than real time, but given the wrong situation could slow down the model where it could not make real time predictions; however, because the results were overall positive, there likely exists a cheaper and more efficient way to direct traffic.
+It does perform faster than real-time, but given the wrong situation, could slow down the model to where it could not make real-time predictions.
+However, because the results were overall positive, there likely exists a cheaper and more efficient way to direct traffic.
 
 While deep Q-Learning with MARL objectively performed well, computation times were relatively slow and implementation would prove very costly.
-Because of this, the MDOT should also investigate sensor based traffic signal controllers.
+Because of this, the MDOT should also investigate sensor-based traffic signal controllers.
 These signals are cheaper to install, but consider timers and congestion levels with more strictly defined rules.
 Implementing these would avoid the high cost of training, implementation, and upkeep.
-Because using state data increased the performance with neural networks, its likely that a simple understanding of the state of the road could help direct traffic without implementing an expensive and unpredictable model.
+Because using state data increased the performance with neural networks, it's likely that a simple understanding of the state of the road could help direct traffic without implementing an expensive and unpredictable model.
 While still having some cost, it would decrease time spent driving and keep the roads of Mississippi safer.
@@
 
-In order to download the code one would need to go to GitHub, and view chandlerweeks/TrafficSignalManagement.
+In order to download the code one would need to go to GitHub and view chandlerweeks/TrafficSignalManagement.
 The link to the repository is: https://github.com/ChandlerWeeks/TrafficSignalManagement.
 Once here, fork the repository and make a local clone on your device.
-If a release is available at the time of reading the paper, that link will allow direct access of the codebase.
+If a release is available at the time of reading the paper, that link will allow direct access to the codebase.
 
@@
 
-Once every prerequisite has been setup, to test installation and configuration, one should open SUMO-GUI and open the file in data/simulation.sumocfg.
+Once every prerequisite has been set up, to test installation and configuration, one should open SUMO-GUI and open the file in data/simulation.sumocfg.
 If the file runs and traffic is observed, then the testing traffic flow scenario has been loaded and the code is ready to run.
@@
 In order to run the benchmarks with the model, one must run the trained model first to get the trip information.
-This can be ran through this command \texttt{python -m traffic\_signal\_nn.eval.run\_model --config traffic\_signal\_nn/config/config\_dqn\_city.ini}, which runs the trained model on the same situation as the previous benchmark, for consistency as mentioned in cross-validation.
+This can be run through this command \texttt{python -m traffic\_signal\_nn.eval.run\_model --config traffic\_signal\_nn/config/config\_dqn\_city.ini}, which runs the trained model on the same situation as the previous benchmark, for consistency as mentioned in cross-validation.
 Just like without the neural network, the queue length is not obtained for the benchmarks, so one will need to run \texttt{python .\textbackslash traffic\_signal\_nn\textbackslash eval\textbackslash get\_model\_queue\_length.py --config .\textbackslash traffic\_signal\_nn\textbackslash config\textbackslash config\_dqn\_city.ini --output .\textbackslash data\textbackslash output\textbackslash queue\_length.csv} in order the get the updated queue lengths.
@@
 
-Setup of the project can be rather time consuming, as mentioned, a video can provided to the reviewers running the script completely setup if necessary.
+Setup of the project can be rather time-consuming, as mentioned, a video can provided to the reviewers running the script completely setup if necessary.
 
@@
 
-Large language models were used in assistance of coding, particularly for debugging and assistance with small code snippets, but never used to generate large parts of a function, or codebits.
-The code is developed by human, but AI assisted with some lines.
+Large language models were used in assistance of coding, particularly for debugging and assistance with small code snippets, but never used to generate large parts of a function or code bits.
+The code is developed by humans, but AI assisted with some lines.
 
